### 2020-07-16 cases_and_policies/rmd/datasummary.Rmd

---
title       : ""
subtitle    : "Data description and summary statistics"
author      : Paul Schrimpf, Hiro Kasahara, Victor Chernozhukov
job         :
date        : "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: "covid.bib"
link-citations: true
output      :
    html_document :
        toc : true
        toc_depth : 3
        toc_float : true
        number_sections : true
        code_folding: hide
        fig_width: 8
        fig_height: 6
## To create html files from this, in R enter `library(rmarkdown);
## rmarkdown::render("regressionsWithPlots.Rmd")`
##

## to create the gh-pages hosted website in R enter
## rmarkdown::render_site()
##
## Then in a shell
## $ ghp-import -p -n -m "$(date)" _site
##

---

```{r setup}
library(ggthemes)
library(ggplot2)
library(gridExtra)
library(knitr)
library(kableExtra)

rootdir <- system("git rev-parse --show-toplevel", intern=TRUE)[1]
datafile <- paste(rootdir,"cases_and_policies/data/covidstates.Rda", sep="/")
cat("Loading data last updated on ", as.character(file.mtime(datafile)), "\n")
cat("Run ", paste(rootdir,"cases_and_policies/R/cases_and_policies.R", sep="/"), " to update data.\n")
load(datafile)
source(paste(rootdir,"cases_and_policies/R/utils.R",sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/varlabels.R", sep="/"))

plottheme <- figtheme
colors <-  scale_color_solarized

kable <- function(tbl, ...) kable_styling(knitr::kable(tbl, ...),
                                          bootstrap_options=c("striped","condensed","responsive"),
                                          full_width=TRUE)

cs <- covidstates
cs <- cs[cs$fips<60, ] # drop territories (Puerto Rico, Guam, etc)
cs <- cs[order(cs$state, cs$date),]
```

# Comparing Datasets on Cases and Deaths

```{r, cache=TRUE}
vars <- c("cases.nyt","cases","positive")
tbl <- cor(cs[,vars],use="complete")
rownames(tbl) <- colnames(tbl) <- c("NYT","JHU","CTP")
kable(tbl, digits=5, caption="Correlation of cumulative cases \\label{tab:casecor}")

tbl2 <- matrix(NA,3,3)
for(i in 1:3) {
  for(j in 1:3) {
    tbl2[i,j] <- mean(cs[,vars[i]]==cs[,vars[j]], na.rm=TRUE)
  }
}
rownames(tbl2) <- rownames(tbl)
colnames(tbl2) <- colnames(tbl)
kable(tbl2, digits=5,
      caption="Portion of days that cumulative cases are equal between data sets \\label{tab:eqcases}")

```

## Daily Cases

```{r newcases, cache=FALSE}
startday <- as.Date("2020-02-15")
cs$newcases.nyt <- paneldiff(cs$cases.nyt, cs$state, cs$date, lag=1)
cs$newcases.jhu <- paneldiff(cs$cases, cs$state, cs$date, lag=1)
cs$newcases.ctp <- paneldiff(cs$positive, cs$state, cs$date, lag=1)

fctp <- ggplot(cs, aes(x=date, y=newcases.ctp, colour=state)) + geom_line(alpha=1) +
  xlim(as.Date(startday),max(cs$date)) + scale_y_log10() + plottheme +
  ggtitle("CTP") + ylab("New cases") +
  theme(legend.position = "none")
fnyt <- ggplot(cs, aes(x=date, y=newcases.nyt, colour=state)) + geom_line(alpha=1) +
  xlim(as.Date(startday),max(cs$date)) + scale_y_log10() + plottheme +
  ggtitle("NYT") + ylab("New cases") +
  theme(legend.position = "none")
fjhu <- ggplot(cs, aes(x=date, y=newcases.jhu, colour=state)) + geom_line(alpha=1) +
  xlim(as.Date(startday),max(cs$date)) + scale_y_log10() + plottheme +
  ggtitle("JHU") + ylab("New cases") +
  theme(legend.position = "none")

grid.arrange(fnyt,fjhu,fctp)

X  <- cs[,c("newcases.nyt","newcases.jhu","newcases.ctp")]
X[X<0] <- NA
X[X==0] <- 0.1
tbl  <- colSums(cs[,c("newcases.nyt","newcases.jhu","newcases.ctp")]<0, na.rm=TRUE)
tbl <- cor(log(X), use="complete")
tbl <- rbind(tbl, diag(cov(log(X), use="complete")))
colnames(tbl) <- c("NYT","JHU","CTP")
rownames(tbl) <- c("NYT","JHU","CTP","Variance")

kable(tbl, digits=4, caption="Correlation and variance of log daily new cases\\label{tab:newcasecor}")
```

## Weekly Cases

```{r}
cs$newcases.nyt <- paneldiff(cs$cases.nyt, cs$state, cs$date, lag=7)
cs$newcases.jhu <- paneldiff(cs$cases, cs$state, cs$date, lag=7)
cs$newcases.ctp <- paneldiff(cs$positive, cs$state, cs$date, lag=7)
X  <- cs[,c("newcases.nyt","newcases.jhu","newcases.ctp")]
X[X<0] <- NA
X[X==0] <- 0.1
tbl  <- colSums(cs[,c("newcases.nyt","newcases.jhu","newcases.ctp")]<0, na.rm=TRUE)
tbl <- cor(log(X), use="complete")
tbl <- rbind(tbl, diag(cov(log(X), use="complete")))
colnames(tbl) <- c("NYT","JHU","CTP")
rownames(tbl) <- c("NYT","JHU","CTP","Variance")
kable(tbl, digits=4, caption="Correlation and variance of log weekly new cases\\label{tab:weekcasecor}")

fctp <- ggplot(cs, aes(x=date, y=newcases.ctp, colour=state)) + geom_line(alpha=1) +
  xlim(startday,max(cs$date)) + scale_y_log10() + plottheme +
  ggtitle("CTP") + ylab("New cases") +
  theme(legend.position = "none")
fnyt <- ggplot(cs, aes(x=date, y=newcases.nyt, colour=state)) + geom_line(alpha=1) +
  xlim(startday,max(cs$date)) + scale_y_log10() + plottheme +
  ggtitle("NYT") + ylab("New cases") +
  theme(legend.position = "none")
fjhu <- ggplot(cs, aes(x=date, y=newcases.jhu, colour=state)) + geom_line(alpha=1) +
  xlim(startday,max(cs$date)) + scale_y_log10() + plottheme +
  ggtitle("JHU") + ylab("New cases") +
  theme(legend.position = "none")

grid.arrange(fnyt,fjhu,fctp)
```

## Deaths

```{r deathplots, cache=TRUE}
cs$newdeaths.nyt <- paneldiff(cs$deaths.nyt, cs$state, cs$date, lag=7)
cs$newdeaths.jhu <- paneldiff(cs$deaths, cs$state, cs$date, lag=7)
cs$newdeaths.ctp <- paneldiff(cs$deaths.ctp, cs$state, cs$date, lag=7)
X  <- cs[,c("newdeaths.nyt","newdeaths.jhu","newdeaths.ctp")]
tbl  <- colSums(cs[,c("newdeaths.nyt","newdeaths.jhu","newdeaths.ctp")]<0, na.rm=TRUE)
tbl <- cor(X, use="complete")
tbl <- rbind(tbl, diag(cov(X, use="complete")))
colnames(tbl) <- c("NYT","JHU","CTP")
rownames(tbl) <- c("NYT","JHU","CTP","Variance")
kable(tbl, digits=4, caption="Correlation and variance of weekly deaths\\label{tab:weekdeathcor}")

fctp <- ggplot(cs, aes(x=date, y=newdeaths.ctp, colour=state)) + geom_line(alpha=1) +
  xlim(startday,max(cs$date)) + scale_y_log10() + plottheme +
  ggtitle("CTP") + ylab("Deaths") +
  theme(legend.position = "none")
fnyt <- ggplot(cs, aes(x=date, y=newdeaths.nyt, colour=state)) + geom_line(alpha=1) +
  xlim(startday,max(cs$date)) + scale_y_log10() + plottheme +
  ggtitle("NYT") + ylab("Deaths") +
  theme(legend.position = "none")
fjhu <- ggplot(cs, aes(x=date, y=newdeaths.jhu, colour=state)) + geom_line(alpha=1) +
  xlim(startday,max(cs$date)) + scale_y_log10() + plottheme +
  ggtitle("JHU") + ylab("Deaths") +
  theme(legend.position = "none")

grid.arrange(fnyt,fjhu,fctp)
```

# Testing data

Our data on tests is from the [Covid Tracking
Project](https://covidtracking.com/). This dataset contains the following variables.

```{r , cache=TRUE}
kable(dataguide[dataguide$source=="Covid-Tracking Project",c("name","source")])
```

We have been using the `totalTestResults` variable.

```{r testplots, cache=TRUE}
fign <- ggplot(aggregate(totalTestResults ~ date, FUN=function(x) sum(!is.na(x)), data=cs)) +
  geom_line(aes(x=date, y=totalTestResults, colour="Tests"), size=1.5) +
  geom_line(data=aggregate(cases ~ date, FUN=function(x) sum(!is.na(x)), data=cs),
            aes(x=date, y=cases, colour="Cases"), size=1.3) +
  geom_line(data=aggregate(deaths ~ date, FUN=function(x) sum(!is.na(x)), data=cs),
            aes(x=date, y=deaths, colour="Deaths"), size=1.3) +
            ylab("Number non-missing") +
            xlab("") + plottheme + ggtitle("Number of non-missing observations") + colors()

p <- c(0.05, 0.25, 0.5, 0.75, 0.95)
adf <- aggregate(totalTestResults ~ date,
                 FUN=function(x) quantile(x, probs=p,
                                          na.rm=TRUE), data=cs)

## The following code sucks and the fact that the equivalent loop
## doesn't work reflects all that is wrong with ggplot2 and the
## tidyverse
p <- as.factor(p)
s <- 1.5
figm <- ggplot(adf) +
  geom_line(aes(x=date, y=adf$totalTestResults[,1], color=p[1]), size=s) +
  geom_line(aes(x=date, y=adf$totalTestResults[,2], color=p[2]), size=s) +
  geom_line(aes(x=date, y=adf$totalTestResults[,3], color=p[3]), size=s) +
  geom_line(aes(x=date, y=adf$totalTestResults[,4], color=p[4]), size=s) +
  geom_line(aes(x=date, y=adf$totalTestResults[,5], color=p[5]), size=s) +
  plottheme + scale_y_log10() + ylab("Cumulative tests") +
  colors() + ggtitle("Quantiles of cumulative tests") +
  geom_line(data=cs, aes(x=date, y=totalTestResults, group=state), alpha=0.2) +
  theme(legend.position="bottom")

grid.arrange(fign, figm, layout_matrix=matrix(1:2, nrow=2), heights=c(0.5,1))
```

The top panel shows that states began reporting test numbers in early
March at earliest, and test numbers are included for all states by mid
March. A few states began reporting cases earlier than March, but
mostly case reporting and test reporting coincide.

In the bottom panel, the thick colored lines are quantiles of total
cumulative tests, and the gray lines are tests in each state.

# Policies

```{r, cache=TRUE}
rvars <- dataguide[dataguide$source=="Raifman et al", "name"]
pols  <- unique(cs[,c("state","fips",rvars)])
di <- sapply(cs[,rvars], function(x) inherits(x,"Date"))
pvars <- rvars[di]

tbl <- cbind(sapply(pols[,pvars], function(x) sum(!is.na(x)), USE.NAMES=FALSE),
             sapply(pols[,pvars], function(x) min(x, na.rm=TRUE), USE.NAMES=FALSE),
             sapply(pols[,pvars], function(x) median(x, na.rm=TRUE), USE.NAMES=FALSE),
             sapply(pols[,pvars], function(x) max(x, na.rm=TRUE), USE.NAMES=FALSE))
tbl[,2:4] <- as.character(as.Date(as.numeric(tbl[,2:4]), origin=as.Date("1970-01-01", format="%Y-%m-%d")))
rownames(tbl) <- gsub("\\."," ",pvars)
colnames(tbl) <- c("N","Min","Median","Max")
kable(tbl, digits=0, caption="State Policies \\label{tab:policies}", )
```

```{r policyplot, cache=TRUE, warning=FALSE}
rvars <- dataguide[dataguide$source=="Raifman et al", "name"]
pols  <- unique(cs[,c("state","fips",rvars)])
di <- sapply(cs[,rvars], function(x) inherits(x,"Date"))
pols <- rvars[di]

plotstatepolicies  <- function(st) {
  sdf  <- subset(cs, state==st)
  x  <- c()
  labels <- c()
  for (lbl in pols) {
    d <- unique(sdf[,lbl])
    i <- which(x==d)
    if (length(i)==0) {
      if (length(x)==0)
        x <- d
      else
        x <- c(x,d)
      labels <- c(labels,lbl)
    } else {
      labels[i]  <- paste(labels[i],lbl,sep=", ")
    }
  }
  labels <- labels[!is.na(x)]
  x <- x[!is.na(x)]
  y <- rep(1, length(x))
  fig <- ggplot(sdf,aes(x=date))  +
    geom_line(aes(y=cases)) +
    geom_line(aes(y=deaths), colour="red") +
    xlim(startday,max(sdf$date)) +
    ylab("count") +
    scale_y_continuous(trans="log10") + plottheme + colors() +
    ggtitle(st)

  fig + annotate("text", x=x,y=y,label=gsub("\\."," ",labels),
                 size=2, angle=90, hjust=0)

}

i <- 0
for (st in unique(cs$state)) {
  if (i == 0) fig <- list()
  fig <- c(fig, list(plotstatepolicies(st)))
  i <- i+1
  if (i %% 3 == 0) {
    grid.arrange(grobs=fig, nrow=3, ncol=1)
    i = 0
  }
}
if (i != 0) {
  fig <- c(fig, rep(list(ggplot() + plottheme), 3-i))
  grid.arrange(grobs=fig, nrow=3, ncol=1)
}
```

```{r npolicies, warning=FALSE, cache=TRUE}
df <- cs[order(cs$state,cs$date),]
pdates <- c("Date.closed.K.12.schools",
            "Stay.at.home..shelter.in.place",
            "Closed.movie.theaters",
            "Closed.restaurants.except.take.out",
            "Closed.non.essential.businesses",
            "Mandate.face.mask.use.by.all.individuals.in.public.spaces",
            "Mandate.face.mask.use.by.employees.in.public.facing.businesses",
            "State.of.emergency")
enddates <- c(NA,
              "End.relax.stay.at.home.shelter.in.place",
              "Reopened.movie.theaters",
              "Reopen.restaurants",
              "Began.to.reopen.businesses.statewide",
              NA,
              NA,
              NA)
stopifnot(length(enddates)==length(pdates))

pvars <- c("pk12", "pshelter", "pmovie", "prestaurant", "pnonessential",
           "pmaskall","pmaskbus", "psoe")
for(i in 1:length(pdates)) df[,pvars[i]] <-
                             smoothedpolicy(df,pdates[i],enddatevar=enddates[i],type="ma",bw=0)

pols <- c("psoe","pmaskbus","pk12", "pshelter", "pmovie", "prestaurant", "pnonessential")

ppplot <- function(pol, df) {
  df$y <- df[,pol]
  adf <- aggregate(y ~ date, data=df, function(x) mean(x, na.rm=TRUE))
  fig <- ggplot(adf, aes(x=date, y=y)) + geom_line(size=1.2) +
    plottheme + ylab("Portion of States") +
    colors() + ggtitle(varlabels[varlabels$Variable==pol,"Label"]) +
    theme(legend.position="none") +
    xlim(startday, max(df$date)) + ylim(0,1)
  return(fig)
}
figs <- lapply(pols,function(p) ppplot(p, df))

for (i in 1:length(pols)) {
  pdf(sprintf("%s/tex/tables_and_figures/%s_p.pdf",rootdir, pols[i]), width=4, height=2)
  print(figs[[i]])
  dev.off()
}
grid.arrange(grobs=figs[2:7], nrow=3, ncol=2)
```

# Social Distancing Measures

```{r, warning=FALSE}
gv <- dataguide$name[dataguide$source=="Google Mobility Reports"]
hv <- dataguide$name[dataguide$source=="Homebase"]
uv <- c("daily_distance_diff", "daily_visitation_diff",
        "encounters_rate", "log_encounters_rate")
cs$log_encounters_rate <- log(1+cs$encounters_rate)
```

## Google Mobility Reports

```{r gplots, cache=TRUE, warning=FALSE}
## fign <-
##   ggplot(aggregate(retail_and_recreation_percent_change_from_baseline
##                    ~ date,
##                    FUN=function(x) sum(!is.na(x)), data=cs)) +
##   geom_line(aes(x=date, y=retail_and_recreation_percent_change_from_baseline), size=1) +
##   ylab("Number non-missing") +
##   theme(legend.position = "none") +
##   xlab("") + plottheme + ggtitle("Number of non-missing observations of Google Mobility Reports") +
##   colors() + xlim(startday, max(cs$date))
## fign

quantileplot <- function(v, p=c(0.05, 0.25, 0.5, 0.75, 0.95), data=cs) {
  data$y <- data[,v]
  adf <- aggregate(y ~ date, data=data,
                   FUN=function(x) quantile(x, probs=p, na.rm=TRUE))
  p <- as.factor(p)
  s <- 1.2
  txt <- data.frame(date=max(adf$date), y=adf$y[nrow(adf$y),], p=p)
  names(adf)[1] <- "date"
  figm <- ggplot(adf) +
    geom_line(aes(x=date, y=adf$y[,1], color=p[1]), size=s) +
    geom_line(aes(x=date, y=adf$y[,2], color=p[2]), size=s) +
    geom_line(aes(x=date, y=adf$y[,3], color=p[3]), size=s) +
    geom_line(aes(x=date, y=adf$y[,4], color=p[4]), size=s) +
    geom_line(aes(x=date, y=adf$y[,5], color=p[5]), size=s) +
    geom_text(data=txt,aes(x=date,y=y,color=p, label=p,
                           hjust="left",vjust="center", size=2)) +
    plottheme + ylab("Percent Change") +
    colors() + ggtitle(TeX(paste("Evolution of",
                             varlabels[varlabels$Variable==v,"Label"]))) +
    geom_line(data=data, aes(x=date, y=y, group=state), alpha=0.2) +
    theme(legend.position="none") +
    #labs(color='Quantile')  +
    xlim(startday, max(data$date))
}

figm <- lapply(gv, quantileplot)

for (i in seq(1,length(figm),by=2)) {
  grid.arrange(grobs=figm[i:(i+1)], nrow=2)
}
for (i in 1:length(figm)) {
  vl <- strsplit(gv[i],"_")[[1]][1]
  pdf(sprintf("%s/tex/tables_and_figures/%s.pdf",rootdir,vl), width=7, height=3.5)
  print(figm[[i]])
  dev.off()
}

```

The following figure shows the residuals from projecting each of the
above variables on the social distancing metrics we use in our main
analysis,
```{r, cache=TRUE}
sdused <- c(#"residential_percent_change_from_baseline",
    "retail_and_recreation_percent_change_from_baseline",
    "transit_stations_percent_change_from_baseline",
    "workplaces_percent_change_from_baseline",
    "grocery_and_pharmacy_percent_change_from_baseline")
kable(t(varlabels[sapply(sdused, function(x) which(x==varlabels$Variable)),"Label"]))
```
no smoothing is used in the figure.

The table shows the $R^2$ from regressing each of the social distance
measures on the ones used in our main regressions. We report this both
with and without smoothing.

```{r, cache=TRUE}
residanalysis <- function(yvars, xvars=sdused, df=cs) {

  madf <- df[,c("state","date",yvars,xvars)]
  for(v in c(yvars, xvars)) {
    madf[,v] <- panelma(madf[,v], madf$state, madf$date, 7)
  }

  r <- lapply(yvars, function(v) {
    if (v %in% xvars) {
      list(r.squared=1, r.squared.ma=1, residuals=rep(0,nrow(df)))
    } else {
      m <- lm(as.formula(paste(v,"~",
                               paste(xvars, collapse=" + "), sep=" ")),
              data=df,na.action=na.exclude)
      ma <- lm(as.formula(paste(v,"~",
                               paste(xvars, collapse=" + "), sep=" ")),
               data=madf,na.action=na.exclude)
      list(r.squared=summary(m)$r.squared,
           r.squared.ma=summary(ma)$r.squared,
           residuals=residuals(m))
    }
  })

  tmp <- data.frame()
  for (i in 1:length(yvars)) {
    tmp <- rbind(tmp,
                 data.frame(var=varlabels[varlabels$Variable==yvars[i],"Label"],
                            date=df$date, state=df$state,
                            residuals=r[[i]]$residuals))
  }
  tbl <- t(matrix(c(sapply(r, function(x) x$r.squared),
                    sapply(r, function(x) x$r.squared.ma)),
                 nrow=length(yvars)))
  colnames(tbl) <- varlabels[sapply(yvars, function(x) which(x==varlabels$Variable)),"Label"]
  rownames(tbl) <- c("No smoothing", "7 day moving average")

  fig <- ggplot(tmp, aes(x=date, y=residuals,color=var)) +
    geom_point(alpha=0.7) + colors() + plottheme + ylab("Residuals") +
    labs(color="") + theme(legend.position=c(0.05, 0.95)) +
    xlim(as.Date("2020-03-01"),max(df$date))
  return(list(fig, tbl))
}

out <- residanalysis(gv, sdused)

kable(out[[2]],
      caption="Variance explained by Google Mobility measures used in regressions")

out[[1]]
```

## Homebase

```{r hplots, cache=TRUE, warning=FALSE}
## fign <-
##   ggplot(aggregate(percentchangehours ~ date,
##                    FUN=function(x) sum(!is.na(x)), data=cs)) +
##   geom_line(aes(x=date, y=percentchangehours), size=1) +
##   ylab("Number non-missing") +
##   theme(legend.position = "none") +
##   xlab("") + plottheme + ggtitle("Number of non-missing observations of HomeBase Variables") +
##   colors() + xlim(startday, max(cs$date))
## fign

figm <- lapply(hv, quantileplot)
grid.arrange(grobs=figm, nrow=2)

out <- residanalysis(hv, sdused)

kable(out[[2]],
      caption="Variance explained by Google Mobility measures used in regressions")

out[[1]]
```

<!-- ## Unacast -->

<!-- ```{r uplots, cache=TRUE, warning=FALSE} -->
<!-- ## fign <- -->
<!-- ##   ggplot(aggregate(encounters_rate ~ date, -->
<!-- ##                    FUN=function(x) sum(!is.na(x)), data=cs)) + -->
<!-- ##   geom_line(aes(x=date, y=encounters_rate), size=1) + -->
<!-- ##   ylab("Number non-missing") + -->
<!-- ##   theme(legend.position = "none") + -->
<!-- ##   xlab("") + plottheme + ggtitle("Number of non-missing observations of Unacast Variables") + -->
<!-- ##   colors() + xlim(startday, max(cs$date)) -->
<!-- ## fign -->

<!-- figm <- lapply(uv, quantileplot) -->
<!-- figm <- lapply(figm, function(x) x + ylab("")) -->
<!-- for (i in seq(1,length(figm), by=2)) { -->
<!--   grid.arrange(grobs=figm[i:(i+1)], nrow=2) -->
<!-- } -->

<!-- out <- residanalysis(uv, sdused) -->

<!-- kable(out[[2]], -->
<!--       caption="Variance explained by Google Mobility measures used in regressions") -->

<!-- out[[1]] -->
<!-- ``` -->

<!-- ## SafeGraph -->

<!-- ```{r splots, cache=TRUE, warning=FALSE} -->
<!-- sv <- c("distance_traveled_from_home", "completely_home_device_count", -->
<!--         "median_home_dwell_time", "part_time_work_behavior_devices", -->
<!--         "full_time_work_behavior_devices","portion_completely_home", -->
<!--         "portion_full_time_work", "portion_part_time_work") -->
<!-- cs$portion_completely_home <- -->
<!--   cs$completely_home_device_count/cs$device_count -->
<!-- cs$portion_full_time_work <- -->
<!--   cs$full_time_work_behavior_devices/cs$device_count -->
<!-- cs$portion_part_time_work <- -->
<!--   cs$part_time_work_behavior_devices/cs$device_count -->

<!-- ## fign <- -->
<!-- ##   ggplot(aggregate(distance_traveled_from_home ~ date, -->
<!-- ##                    FUN=function(x) sum(!is.na(x)), data=cs)) + -->
<!-- ##   geom_line(aes(x=date, y=distance_traveled_from_home), size=1) + -->
<!-- ##   ylab("Number non-missing") + -->
<!-- ##   theme(legend.position = "none") + -->
<!-- ##   xlab("") + plottheme + -->
<!-- ##   ggtitle("Number of non-missing observations of Safegraph Variables") + -->
<!-- ##   colors() + xlim(startday, max(cs$date)) -->
<!-- ## fign -->

<!-- figm <- lapply(sv, quantileplot) -->
<!-- figm <- lapply(figm, function(x) x + ylab("")) -->
<!-- for (i in seq(1,length(figm), by=2)) { -->
<!--   grid.arrange(grobs=figm[i:(i+1)], nrow=2) -->
<!-- } -->

<!-- out <- residanalysis(sv, sdused) -->

<!-- kable(out[[2]], -->
<!--       caption="Variance explained by Google Mobility measures used in regressions") -->

<!-- out[[1]] -->
<!-- ``` -->

# Variables

```{r varlabels, cache=TRUE}
source(paste(rootdir,"cases_and_policies/rmd/varlabels.R", sep="/"))
kable(varlabels)
```

# Case and death growth

```{r cdgrowth, cache=FALSE}
source(paste(rootdir,"cases_and_policies/R/dataprep.R",sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/varlabels.R", sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/regprep.R", sep="/"))
fig <- quantileplot("dlogdc", data=df) + ylab("") + xlim(c(as.Date("2020-03-01"),max(df$date+3)))
pdf(sprintf("%s/tex/tables_and_figures/casequantiles.pdf",rootdir), width=7, height=3.5)
print(fig)
dev.off()

fig <- quantileplot("dlogdd", data=df) + ylab("") + xlim(c(as.Date("2020-03-01"),max(df$date+3)))
pdf(sprintf("%s/tex/tables_and_figures/deathquantiles.pdf",rootdir), width=7, height=3.5)
print(fig)
dev.off()

```