### 2020-07-23 cases_and_policies/rmd/counterfactuals.Rmd
Reorganized counterfactuals code.
Added longer description of counterfactuals.

Added counterfactual with fixed residuals for masks with cases.

---
title       : "Causal impact of masks, policies, behavior on early COVID-19 pandemic in the U.S."
subtitle    : "Counterfactuals"
author      :  Victor Chernozhukov, Paul Schrimpf, Hiro Kasahara
job         :
date        : "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: "covid.bib"
link-citations: true
always_allow_html: true
output      :
    html_document :
        toc : true
        toc_depth : 3
        toc_float : true
        number_sections : true
        #theme : journal
        #css : 628notes.css
        code_folding: hide
        #lib_dir : deps
        self_cononontained : true
        fig_width: 8
        fig_height: 6
---

 <!-- To create html files from this, in R enter  -->
 <!-- library(rmarkdown) -->
 <!-- rmarkdown::render("regressionsWithPlots.Rmd") -->
 <!-- To create tex files from this, in R enter ` -->
 <!-- library(rmarkdown) -->
 <!-- rmarkdown::render("regressionsWithPlots.Rmd", output_format=latex_fragment()) -->

 <!-- to create the gh-pages hosted website in R enter -->
 <!-- rmarkdown::render_site() -->

 <!-- Then in a shell  -->
 <!-- $ ghp-import -p -n -m "$(date)" _site -->

```{r, include=FALSE}
options(knitr.table.format = function() {
  if (knitr::is_latex_output()) 'latex' else 'pandoc'
})
if (knitr::is_latex_output()) {
  knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo=FALSE)
}
```

# Data Preparation
```{r setup, cache=FALSE, warning=FALSE, include=FALSE}

library(lfe)
library(stargazer)
library(knitr)
library(plm)
library(latex2exp)
library(ggplot2)
library(ggthemes)
library(estimatr)
library(gridExtra)
library(grid)
library(mvtnorm)
library(kableExtra)
library(data.table)
library(reshape2)

sgtype <- opts_knit$get("rmarkdown.pandoc.to")
sgstyle <- 'default'
colors <-  scale_color_solarized
colors_fill <- scale_fill_solarized
rootdir <- system("git rev-parse --show-toplevel", intern=TRUE)[1]
source(paste(rootdir,"cases_and_policies/R/dataprep.R",sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/varlabels.R", sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/regprep.R", sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/generatetables.R", sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/bootstrap_felm.R", sep="/"))

datelims <- c(as.Date("2020-03-07"), max(df$date))
```

# Counterfactuals

```{r cffuncs, cache=FALSE}
source(paste(rootdir,"cases_and_policies/rmd/counterfactuals.R", sep="/"))
```

```{r statesim, cache=FALSE}
infovars <- list(c("dlogdc", "logdc")) #,
#c("dlogdc", "logdc"))
#infovars2 <- list(c( "logdc"))
tvars <- "dlogtests"
xlist <-list("")
interactions <-list("month", statevars)
ilist <- list(interactions) #, interactions)
iv <- list("0") #,"(testratedc ~ lag(testratedc,14))")

L <- 14
sdf <- subset(df, df$date>=datelims[1]) #+L)
regs <- mainregressions(sdf,
                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
m <- regs$piy[[1]]$reg
cnames <- c(pols, infovars[[1]])
di <- dieff_table(lapply(regs$pib[[1]][1:4], function(x) x$reg$coef[,1]),
                  regs$pbiy[[1]]$reg$coef[,1],
                  regs$piy[[1]]$reg$coef[,1],
                  policies=cnames)

pib <- lapply(regs$pib[[1]][1:4], function(x) x$reg)
pbiy <- regs$pbiy[[1]]$reg
piy <- regs$piy[[1]]$reg

S <- 999
bs <- bootstrap_felm(sdf, c(pib,list(pbiy,piy)), S=S)
bdi <- lapply(1:S, function(i)
  dieff_table(lapply(1:4, function(j) bs[[j]][,i]),
              bs[[5]][,i], bs[[6]][,i],
              policies=cnames)
  )

subcoefs <- function(di, coef) {
  for (r in 1:(nrow(di)-1)) {
    i <- grep(rownames(di)[r] , rownames(coef))
    coef[i,1] <- di[r,"Average"]
  }
  return(coef)
}

bcoefs <- t(do.call(cbind, lapply(1:S, function(i) {
  coef <- matrix(bs[[6]][,i],ncol=1)
  rownames(coef) <- rownames(bs[[6]])
  return(subcoefs(bdi[[i]],coef ))
})))

V <- cov(bcoefs)
m$coefficients <- subcoefs(di, m$coefficients)
m$vcv <- V
m$clustervcv <- V
m$robustvcv <- V

simmodel <- createforwardsim(m, sdf)

# remove policies
cf <- sdf
cf[,pols] <- 0
nopmodel <- createforwardsim(m, cf)

rerunsim=FALSE
if (!file.exists("cfnop.Rda") || rerunsim) {
  cfnoplist <- lapply(unique(sdf$state), function(s)
    nopstatesim(s,simobs=simmodel,simcf=nopmodel,data=sdf))
  allmeans <- rbindlist(lapply(cfnoplist, getmeans))
  save(cfnoplist,allmeans, file="cfnop.Rda")
}
load("cfnop.Rda")
cfnop <- do.call(rbind, lapply(cfnoplist, function(x) x$edf))
```

### Removing Policies

We now examine the impact changing from the observed policies to
none. For illustrative purpose, we being by focusing on
Washington. Figure \ref{fig:WA-nop-dlogdc} shows the observed,
estimated average, and counterfactual without policies average of
$\Delta \log \Delta C$ in Washington.

To compute the estimated and counterfactual paths we use the estimates
from column 2 of table \ref{tab:policy-cases}. We set initial $\Delta
\log \Delta C$ and $\log \Delta C$ to their values first observed in
the state we are simulating. We hold all other regressors at their
observed values. Error terms are drawn with replacement from the
residuals. We do this many times and report the average over draws of
the residuals. To calculate confidence intervals, we repeat the above
with coefficients drawn randomly from their asymptotic
distribution. The shaded region is a pointwise 90\% confidence
interval.

To be more precise, we can rewrite our regression model as:

$$
  Y_{it} = P_{it}'\alpha + \sum_{\ell=1}^L \rho_\ell Y_{it-\ell} + X_{it}'\beta + \epsilon_{it}
$$

where $Y_{it} = \log \Delta C_{it}$, $P_{it}$ are policies, and
$X_{it}$ are controls. Each counterfactual involves changing $P_{it}$
to some other value, $P_{it}^\ast$. We then generate counterfactual paths
for cases by in state $i$ conditional on
$Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast \}_{t=1}^T, \theta=(\alpha, \beta, \rho)$ by setting:

$$
  Y_{it}^{\ast,r} = {P_{it}^\ast}'\alpha + \sum_{\ell=1}^L \rho_\ell Y_{it-\ell} + X_{it}'\beta + \hat{\epsilon}_{it}^{\ast,r}
$$

where $\hat{\epsilon}_{it}^{\ast,r}$ are resampled with replacement
from $\{\hat{\epsilon}_{it}\}_{i=1,t=1}^{N,T}$. We do resample the
residuals $R$ times, repeating the above each time. We then compute

$$
\hat{E}[Y_{it}^\ast | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast\}_{t=1}^T, \theta] =
\frac{1}{R} \sum_{r=1}^R Y_{it}^{\ast,r}
$$

Finally, we repeatedly draw $\theta$ from its asymptotic distribution
and repeat all of the above to compute confidence intervals.

For case growth, the same procedure is used, but with $\Delta Y_{it} = \Delta \log \Delta C_{it}$ in place of $Y_{it}$. The top panel of the following figure shows

$$
E[\Delta Y_{it} | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}\}_{t=1}^T]
$$

and

$$
E[\Delta Y_{it}^\ast | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast\}_{t=1}^T]
$$

along with the observed $Y_{it}$, and the bottom panel shows the difference.

```{r WA-nop-dlogdc, cache=TRUE,fig.cap="Case growth without policies in Washington \\label{fig:WA-nop-dlogdc}"}
st <- "Washington"
f <- cfplots(st, cfnop, data=df)
grid.arrange(f$figd, f$figdp, nrow=2, top =
                                        textGrob(st,gp=gpar(fontsize=20,font=2)))

```

To look at cases we instead compute,

$$
\hat{E}[\exp(Y_{it}^\ast) | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast\}_{t=1}^T, \theta] =
\frac{1}{R} \sum_{r=1}^R \exp(Y_{it}^{\ast,r})
$$

```{r WA-nop-cases, cache=TRUE, fig.cap="Cases without policies in Washington \\label{fig:WA-nop-cases}"}
grid.arrange(f$figl, f$figp, nrow=2) #,
#top = textGrob(st,gp=gpar(fontsize=20,font=2)))
```

```{r otherstates, cache=TRUE, fig.cap="Effect of Removing Policies"}
if (!knitr::is_latex_output()) {
  for( st in c("New York", "Florida", "Michigan", "Georgia", "Massachusetts")) {
    f <- cfplots(st, cfnop, data=df)
    grid.arrange(f$figd, f$figdp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
    grid.arrange(f$figl, f$figp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
  }
}
```

```{r nop-state-figures, cache=TRUE, warning=FALSE}
states <- c("Washington","Illinois","Massachusetts")
for(st in states) {
  f <- cfplots(st, cfnop, data=df)

  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dgrowth_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-growth_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figd + ylim(c(-1,5)))
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-cases_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figl)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dcases_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}
```

## National

These figures show the average effect on $\Delta \log \Delta C$ across
states, and the total national effect on cases in past week.

That is for the first figure, we take:

$$
\hat{E}[\Delta Y_{it}^\ast | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast\}_{t=1}^T, \theta] =
\frac{1}{N} \sum_{i=1}^{N} \frac{1}{R} \sum_{r=1}^R \Delta Y_{it}^{\ast,r}
$$

and plot the difference of the above between two different values of $P_{it}^\ast$

The second figure shows the total change in cases, that is the change in:

$$
\hat{E}[\exp(Y_{it}^\ast) | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast\}_{t=1}^T, \theta] =
\sum_{i=1}^{N} \frac{1}{R} \sum_{r=1}^R \exp(Y_{it}^{\ast,r}).
$$

The final figure shows
$$
\frac{\sum_{i=1}^{N} \frac{1}{R} \sum_{r=1}^R \exp(Y_{it}^{\ast,r}) - \sum_{i=1}^{N} \frac{1}{R} \sum_{r=1}^R \exp(Y_{it}^r)} {\sum_{i=1}^{N} \frac{1}{R} \sum_{r=1}^R \exp(Y_{it}^{r})}
$$
that is, the (average over residual draws) change in national cases relative to (average over residual draws) national cases.

```{r nop-all, cache=TRUE}
abbrv <- unique(df[,c("state","ST")])
cfnop <- merge(cfnop, abbrv, by="state")

figs <- nationalplots(allmeans,cfnop, "removing policies")

figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dgrowth_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dcases_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nop-rel_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

## Non-Essential Business Closures

```{r business_sim, cache=FALSE}
# remove non essential business closure
cf <- sdf
cf[,"pnonessential"] <- 0
nobmodel <- createforwardsim(m, cf)

if (!file.exists("cfnb.Rda") || rerunsim) {
  cfnblist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=nobmodel, data=sdf))
  nball <- rbindlist(lapply(cfnblist, getmeans))
  save(cfnblist,nball, file="cfnb.Rda")
}
load("cfnb.Rda")
cfnb <- do.call(rbind, lapply(cfnblist, function(x) x$edf))
```

```{r, cache=TRUE}
cfnb <- merge(cfnb, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfnb, data=df)
  f$figp + ggtitle(sprintf("Effect of never shutting down non-essential businesses in %s",st))
  #grid.arrange(f$figd + ggtitle(), f$figdp + ggtitle(), nrow=2, top =
  #textGrob(st,gp=gpar(fontsize=20,font=2)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-nb-dgrowth_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-nb-dcases_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}

figs <- nationalplots(nball, cfnb,
                      "never shutting down non-essential\nbusinesses")

figdev(paste(rootdir,"tex/tables_and_figures/us-nb-dgrowth_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nb-dcases_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nb-rel_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

## Masks

We now look at what happens when all states impose a mask mandate on
April 1st.

```{r mask_sim, cache=FALSE}
cf <- df
cf[,"Mandate.face.mask.use.by.employees.in.public.facing.businesses"]  <- as.Date("2020-04-01")
cf[,"pmaskbus"] <-
  smoothedpolicy(cf,"Mandate.face.mask.use.by.employees.in.public.facing.businesses",type="ma",bw=6)
cf <- subset(cf, cf$date>=datelims[1]) #+L)
maskmodel <- createforwardsim(m, cf)

if (!file.exists("cfmask.Rda") || rerunsim) {
  cfmasklist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=maskmodel, data=sdf))
  maskall <- rbindlist(lapply(cfmasklist, getmeans))
  save(cfmasklist, maskall, file="cfmask.Rda")
}
load("cfmask.Rda")
cfmask <- do.call(rbind, lapply(cfmasklist, function(x) x$edf))
```

```{r, cache=TRUE}
cfmask <- merge(cfmask, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfmask, data=df)
  print(f$figp +  ggtitle(sprintf("Effect of mandating masks on April 1st in %s",st)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dgrowth_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dcases_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}

figs <- nationalplots(maskall, cfmask, "mandating masks on April 1st\n")

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dgrowth_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dcases_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-rel_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

### Holding Residuals fixed

In this subsection, instead of resampling residuals, we condition on them. That is we look at

$$
  Y_{it}^{\ast} = {P_{it}^\ast}'\alpha + \sum_{\ell=1}^L \rho_\ell Y_{it-\ell} + X_{it}'\beta + \hat{\epsilon}_{it}
$$

instead of

$$
  \frac{1}{R} \sum_{r=1}^R Y_{it}^{\ast,r}.
$$

```{r masks-resids-fixed}
simmodel1 <- function(st, samplecoefs=FALSE, nsim=1)
  simmodel(st, samplecoefs=samplecoefs, nsim=nsim,
           shockgen=NULL)
maskmodel1 <- function(st, samplecoefs=FALSE, nsim=1)
  maskmodel(st, samplecoefs=samplecoefs, nsim=nsim,
            shockgen=NULL)

rerunsim <- FALSE
if (!file.exists("cfmask1.Rda") || rerunsim) {
  cfmasklist1 <- lapply(unique(df$state),
                        function(s) nopstatesim(s,simobs=simmodel1,simcf=maskmodel1, data=sdf, nresid=1))
  maskall1 <- rbindlist(lapply(cfmasklist1, getmeans))
  save(cfmasklist1, maskall1, file="cfmask1.Rda")
}
load("cfmask1.Rda")
cfmask1 <- do.call(rbind, lapply(cfmasklist1, function(x) x$edf))

for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfmask1, yvar="deaths", data=df)
  print(f$figp +  ggtitle(sprintf("Effect of mandating masks on April 1st in %s",st)))

}

figs1 <- nationalplots(maskall1, cfmask1, "mandating masks on April 1st\n")

for (f in figs1) print(f)
```

## Stay at home

We now look at what happens if stay at home orders were never issued.

```{r home, cache=FALSE}
cf <- df
cf[,"pshelter"] <- 0
cf <- subset(cf, cf$date>=datelims[1]) #+L)
noshelter <- createforwardsim(m, cf)

if (!file.exists("cfshelter.Rda") || rerunsim) {
  cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf))
  shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
  save(cfshelterlist, shelterall, file="cfshelter.Rda")
}
load("cfshelter.Rda")
cfshelter <- do.call(rbind, lapply(cfshelterlist, function(x) x$edf))
```

```{r, cache=TRUE}
cfshelter <- merge(cfshelter, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfshelter, data=df)
  print(f$figp +  ggtitle(sprintf("Effect of no stay at home order in %s",st)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dgrowth_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dcases_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}

figs <- nationalplots(shelterall, cfshelter, "no stay at home orders")

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dgrowth_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dcases_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-rel_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

# Deaths

```{r statesim-deaths, cache=FALSE}
infovars <- list(c("dlogdd", "logdd")) #,
#c("dlogdc", "logdc"))
#infovars2 <- list(c( "logdc"))
tvars <- NULL
xlist <-list("")
interactions <-list("month", statevars)
ilist <- list(interactions) #, interactions)
iv <- list("0") #,"(testratedc ~ lag(testratedc,14))")

L <- 21
sdf <- subset(df, df$date>=datelims[1]) #+L)
regs <- mainregressions(sdf,
                        "dlogdd", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)

m <- regs$piy[[1]]$reg

cnames <- c(pols, infovars[[1]])
di <- dieff_table(lapply(regs$pib[[1]][1:4], function(x) x$reg$coef[,1]),
                  regs$pbiy[[1]]$reg$coef[,1],
                  regs$piy[[1]]$reg$coef[,1],
                  policies=cnames)

pib <- lapply(regs$pib[[1]][1:4], function(x) x$reg)
pbiy <- regs$pbiy[[1]]$reg
piy <- regs$piy[[1]]$reg

S <- 999
bs <- bootstrap_felm(sdf, c(pib,list(pbiy,piy)), S=S)
bdi <- lapply(1:S, function(i)
  dieff_table(lapply(1:4, function(j) bs[[j]][,i]),
              bs[[5]][,i], bs[[6]][,i],
              policies=cnames)
  )

bcoefs <- t(do.call(cbind, lapply(1:S, function(i) {
  coef <- matrix(bs[[6]][,i],ncol=1)
  rownames(coef) <- rownames(bs[[6]])
  return(subcoefs(bdi[[i]],coef ))
})))
V <- cov(bcoefs)
m$coefficients <- subcoefs(di, m$coefficients)
m$vcv <- V
m$clustervcv <- V
m$robustvcv <- V

simmodel <- createforwardsim(m, sdf, dvar="dlogdd", var="logdd")

# remove policies
cf <- sdf
cf[,pols] <- 0
nopmodel <- createforwardsim(m, cf, dvar="dlogdd", var="logdd")

rerunsim=FALSE
if (!file.exists("cfnop_deaths.Rda") || rerunsim) {
  cfnoplist <- lapply(unique(df$state), function(s)
    nopstatesim(s,simobs=simmodel,simcf=nopmodel, data=sdf, var="logdd", dvar="dlogdd"))
  allmeans <- rbindlist(lapply(cfnoplist, getmeans))
  save(cfnoplist,allmeans, file="cfnop_deaths.Rda")
}
load("cfnop_deaths.Rda")
cfnop <- do.call(rbind, lapply(cfnoplist, function(x) x$edf))
```

### Removing Policies

```{r WA-nop-dlogdd, cache=TRUE,fig.cap="Case growth without policies in Washington \\label{fig:WA-nop-dlogdc}"}
st <- "Washington"
f <- cfplots(st, cfnop, yvar="deaths", data=df)
grid.arrange(f$figd, f$figdp, nrow=2, top =
                                        textGrob(st,gp=gpar(fontsize=20,font=2)))

```

```{r WA-nop-deaths, cache=TRUE, fig.cap="Cases without policies in Washington \\label{fig:WA-nop-cases}"}
grid.arrange(f$figl, f$figp, nrow=2) #,
#top = textGrob(st,gp=gpar(fontsize=20,font=2)))
```

```{r otherstates-deaths, cache=TRUE, fig.cap="Effect of Removing Policies"}
if (!knitr::is_latex_output()) {
  for( st in c("New York", "Florida", "Michigan", "Georgia", "Massachusetts")) {
    f <- cfplots(st, cfnop, yvar="deaths", data=df)
    grid.arrange(f$figd, f$figdp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
    grid.arrange(f$figl, f$figp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
  }
}
```

```{r nop-state-figures-deaths, cache=TRUE, warning=FALSE}
states <- c("Washington","Illinois","Massachusetts")
for(st in states) {
  f <- cfplots(st, cfnop, yvar="deaths", data=df)

  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dgrowth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-growth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figd + ylim(c(-1,5)))
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-cases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figl)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dcases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}
```

## National

```{r nop-all-deaths, cache=TRUE}
abbrv <- unique(df[,c("state","ST")])
cfnop <- merge(cfnop, abbrv, by="state")

figs <- nationalplots(allmeans,cfnop, "removing policies", yvar="deaths")

figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dgrowth_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dcases_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nop-rel_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

## Non Essential Business Closures

```{r business_sim-deaths, cache=FALSE}
# remove non essential business closure
cf <- sdf
cf[,"pnonessential"] <- 0
nobmodel <- createforwardsim(m, cf, dvar="dlogdd", var="logdd")

if (!file.exists("cfnb_deaths.Rda") || rerunsim) {
  cfnblist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=nobmodel, data=sdf, var="logdd", dvar="dlogdd"))
  nball <- rbindlist(lapply(cfnblist, getmeans))
  save(cfnblist,nball, file="cfnb_deaths.Rda")
}
load("cfnb_deaths.Rda")
cfnb <- do.call(rbind, lapply(cfnblist, function(x) x$edf))
```

```{r, cache=TRUE}
cfnb <- merge(cfnb, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfnb, yvar="deaths", data=df)
  f$figp + ggtitle(sprintf("Effect of never shutting down non-essential businesses in %s",st))
  #grid.arrange(f$figd + ggtitle(), f$figdp + ggtitle(), nrow=2, top =
  #textGrob(st,gp=gpar(fontsize=20,font=2)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-nb-dgrowth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-nb-dcases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}

figs <- nationalplots(nball, cfnb,
                      "never shutting down non-essential\nbusinesses", yvar="deaths")

figdev(paste(rootdir,"tex/tables_and_figures/us-nb-dgrowth_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nb-dcases_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nb-rel_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

## Masks

We now look at what happens when all states impose a mask mandate on
April 1st.

```{r mask_sim-deaths, cache=FALSE}
cf <- df
cf[,"Mandate.face.mask.use.by.employees.in.public.facing.businesses"]  <- as.Date("2020-04-01")
cf[,"pmaskbus"] <-
  smoothedpolicy(cf,"Mandate.face.mask.use.by.employees.in.public.facing.businesses",type="ma",bw=6)
cf <- subset(cf, cf$date>=datelims[1]) #+L)
maskmodel <- createforwardsim(m, cf, dvar="dlogdd", var="logdd")

if (!file.exists("cfmask_deaths.Rda") || rerunsim) {
  cfmasklist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=maskmodel, data=sdf, var="logdd", dvar="dlogdd"))
  maskall <- rbindlist(lapply(cfmasklist, getmeans))
  save(cfmasklist, maskall, file="cfmask_deaths.Rda")
}
load("cfmask_deaths.Rda")
cfmask <- do.call(rbind, lapply(cfmasklist, function(x) x$edf))
```

```{r, cache=TRUE}
cfmask <- merge(cfmask, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfmask, yvar="deaths", data=df)
  print(f$figp +  ggtitle(sprintf("Effect of mandating masks on April 1st in %s",st)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dgrowth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dcases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}

figs <- nationalplots(maskall, cfmask, "mandating masks on April 1st\n", yvar="deaths")

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dgrowth_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dcases_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-rel_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

## Stay at home

We now look at what happens if stay at home orders were never issued.

```{r home-deaths, cache=FALSE}
cf <- df
cf[,"pshelter"] <- 0
cf <- subset(cf, cf$date>=datelims[1]) #+L)
noshelter <- createforwardsim(m, cf, dvar="dlogdd", var="logdd")

if (!file.exists("cfshelter_deaths.Rda") || rerunsim) {
  cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf))
  shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
  save(cfshelterlist, shelterall, file="cfshelter_deaths.Rda")
}
load("cfshelter_deaths.Rda")
cfshelter <- do.call(rbind, lapply(cfshelterlist, function(x) x$edf))
```

```{r, cache=TRUE}
cfshelter <- merge(cfshelter, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfshelter, data=df, yvar="deaths")
  print(f$figp +  ggtitle(sprintf("Effect of no stay at home order in %s",st)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dgrowth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dcases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}

figs <- nationalplots(shelterall, cfshelter, "no stay at home orders", yvar="deaths")

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dgrowth_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dcases_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-rel_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```