### 2020-08-11 cases_and_policies/rmd/regressionsWithPlots.Rmd

---
title       : "Causal impact of masks, policies, behavior on early COVID-19 pandemic in the U.S."
subtitle    : "Tables and Figures"
author      :  Victor Chernozhukov, Paul Schrimpf, Hiro Kasahara
job         :
date        : "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: "covid.bib"
link-citations: true
always_allow_html: true
output      :
    html_document :
        toc : true
        toc_depth : 3
        toc_float : true
        number_sections : true
        #theme : journal
        #css : 628notes.css
        code_folding: hide
        #lib_dir : deps
        self_cononontained : true
        fig_width: 8
        fig_height: 6
---

 <!-- To create html files from this, in R enter  -->
 <!-- library(rmarkdown) -->
 <!-- rmarkdown::render("regressionsWithPlots.Rmd") -->
 <!-- To create tex files from this, in R enter ` -->
 <!-- library(rmarkdown) -->
 <!-- rmarkdown::render("regressionsWithPlots.Rmd", output_format=latex_fragment()) -->

 <!-- to create the gh-pages hosted website in R enter -->
 <!-- rmarkdown::render_site() -->

 <!-- Then in a shell  -->
 <!-- $ ghp-import -p -n -m "$(date)" _site -->

```{r, include=FALSE}
options(knitr.table.format = function() {
  if (knitr::is_latex_output()) 'latex' else 'pandoc'
})
if (knitr::is_latex_output()) {
  knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo=FALSE)
}
```

This file creates the figures and tables in Chernozhukov, Kasahara and
Schrimpf (2020) "Causal Impact of Masks, Policies, Behavior on Early
Covid-19 Pandemic in the U.S."

# Data Preparation
```{r setup, cache=FALSE, warning=FALSE, include=FALSE}

library(lfe)
library(stargazer)
library(knitr)
library(plm)
library(latex2exp)
library(ggplot2)
library(ggthemes)
library(estimatr)
library(gridExtra)
library(grid)
library(mvtnorm)
library(kableExtra)
library(data.table)
library(reshape2)
library(AER)
library(hdm)
library(randomForest)

sgtype <- opts_knit$get("rmarkdown.pandoc.to")
sgstyle <- 'default'
colors <-  scale_color_solarized
colors_fill <- scale_fill_solarized
rootdir <- system("git rev-parse --show-toplevel", intern=TRUE)[1]
source(paste(rootdir,"cases_and_policies/R/exploratory.R", sep="/"))
source(paste(rootdir,"cases_and_policies/R/dataprep.R",sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/varlabels.R", sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/regprep.R", sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/generatetables.R", sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/bootstrap_felm.R", sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/rlassoPanel.R", sep="/"))
```

# Empirical Results

## Correlations

```{r corr, cache=FALSE}

scmat <-
cmat <-  cor(df[,c(bvars,pols,"pindex")], use="pairwise.complete.obs")
  matrix(sprintf("%.2f",cmat),nrow=(length(bvars)+length(pols)+1))

rownames(scmat) <- getlabel(rownames(cmat))
colnames(scmat) <- getlabel(colnames(cmat))
scmat[upper.tri(cmat)] <- ""

ltbl <- knitr::kable(scmat, format="latex", booktabs=TRUE, escape=FALSE,
              align=(rep("c",ncol(scmat)))) %>%
  row_spec(0, angle = 90)
cat(ltbl, file=paste(rootdir,"tex/tables_and_figures/corr.tex",sep="/"))

knitr::kable(scmat)
```

## Test Smoothing

We rearrange and linearly interpolate constant test counts, but do not
apply any additional smoothing.

```{r testsmooth, cache=FALSE}
rearrange <- function(x, id) {
  for (i in unique(id)) {
    xi <- x[id==i]
    xi[!is.na(xi)] <- sort(xi[!is.na(xi)])
    x[id==i] <- xi
  }
  return(x)
}

dropnotincreasing <- function(x, id, t) {
  const <- (x <= panellag(x, id, t))
  const[is.na(const)] <- FALSE
  x[const & x>0] <- NA
  return(x)
}

lininterp <- function(x, id, t) {
  for (i in unique(id)) {
    xi <- x[id==i]
    ti <- t[id==i]
    xi[is.na(xi)] <- approx(ti[!is.na(xi)], xi[!is.na(xi)], ti[is.na(xi)])$y
    x[id==i] <- xi
  }
  return(x)
}

smoothtests <- function(x, id, t) {
  x[is.na(x)] <- 0
  return(lininterp(dropnotincreasing(rearrange(x,id),id,t),id,t))
}

df$neg.interp <- smoothtests(df$negative, df$state, df$date)
df$pos.interp <- df$cases #smoothtests(df$positive, df$state, df$date)
df$tot.interp <- df$neg.interp + df$pos.interp

L <- 7
df$tests <- paneldiff(df$tot.interp, df$state, df$date, lag=L)
df$testrate  <- df$tests/df$Population.2018*1000
df$testrate[df$testrate<0] <- NA
df$dlogtests <- paneldiff(log(sapply(df$testrate, function(x)
  max(x,exp(-1)))), df$state, df$date, lag=L)
df$testratedc <- df$ddeath/df$dcases*df$testrate
df$testratedc[df$dcases==0] <- 0

df$tests.ma <- paneldiff(panelma(df$neg.interp, df$state, df$date, 7) + df$cases,
                         df$state, df$date, lag=L)
df$testrate.ma  <- df$tests.ma/df$Population.2018*1000
df$testrate.ma[df$testrate.ma<0] <- NA
df$dlogtests.ma <- paneldiff(log(sapply(df$testrate.ma, function(x)
  max(x,exp(-1)))), df$state, df$date, lag=L)
df$testratedc.ma <- df$ddeath/df$dcases*df$testrate.ma
df$testratedc.ma[df$dcases==0] <- 0
```

## Cases

```{r caseregs, cache=FALSE, warning=FALSE, results='hide'}
infovars <- list(c("dlogdc", "logdc"),
                 c("dlogdc", "logdc","dlogdc.national", "logdc.national")) #,
#c("dlogdc", "logdc"))
#infovars2 <- list(c( "logdc"))
tvars <- "dlogtests"
xlist <-list("", "")
interactions <-list("month", statevars)
ilist <- list(interactions, interactions)
iv <- list("0","0") #,"(testratedc ~ lag(testratedc,14))")

L <- 14
sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
regs <- mainregressions(sdf,
                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
```

```{r casetables, cache=FALSE, warning=FALSE, results='asis'}
showhtmltables(regs$pib[[1]], regs$pbiy, regs$piy, regs$ip)
savetextables(regs$pib[[1]], NULL, NULL, prefix="behavior")
savetextables(c(regs$pib[[1]][1:4], regs$pib[[2]][1:4]), NULL, NULL, prefix="behavior-NI")

savetextables(c(regs$pib[[1]][1:4]), NULL, NULL, prefix="behavior-cases-stateinfo")
savetextables(c(regs$pib[[2]][1:4]), NULL, NULL, prefix="behavior-cases-nationinfo")
savetextables(NULL, regs$pbiy, regs$piy, prefix="cases")
savetextables(NULL, regs$pbiy[c(1,3)], regs$piy[c(1,3)], prefix="cases-nosplit")
```

```{r casedieff, eval = TRUE, cache=FALSE, warning=FALSE, results='asis'}
for (use.national in c(TRUE,FALSE)) {
  sb <- ifelse(use.national, 2, 1)
  sy <- ifelse(use.national, 3, 1)
  pib <- lapply(regs$pib[[sb]][1:4], function(x) x$reg)
  pbiy <- regs$pbiy[[sy]]$reg
  piy <- regs$piy[[sy]]$reg

  S <- 999
  bs <- bootstrap_felm(sdf, c(pib,list(pbiy,piy)), S=S)

  addse <- function(di, bdi) {
    se <- di*NA
    for (i in 1:length(se)) {
      se[i] <- sd(sapply(bdi, function(b) b[i]))
    }

    tbl <- matrix(NA, nrow=2*nrow(di), ncol=ncol(di))
    for (i in 1:nrow(di)) {
      tbl[2*(i-1) + 1,] <- sapply(1:ncol(di), function(j) printstars(di[i,j], se[i,j]))
      tbl[2*i,] <- sprintf("(%.3f)", se[i,])
    }
    colnames(tbl) <- colnames(di)
    rownames(tbl) <- rep("", nrow(tbl))
    rownames(tbl)[seq(1,nrow(tbl),by=2)] <- relabel(rownames(di))
  return(tbl)
  }

  diall <- dieff_table(lapply(regs$pib[[sb]][1:4], function(x) x$reg$coef[,1]),
                       regs$pbiy[[sy]]$reg$coef[,1],
                       regs$piy[[sy]]$reg$coef[,1],
                       policies=c(pols, infovars[[sb]]), nsum=length(pols))

  bdiall <- lapply(1:S, function(i)
    dieff_table(lapply(1:4, function(j) bs[[j]][,i]),
                bs[[5]][,i], bs[[6]][,i],
                policies=c(pols, infovars[[sb]]), nsum=length(pols))
    )

  tbl <- addse(diall, bdiall)

  #tbl <- tbl[-c(nrow(tbl), nrow(tbl)-1),]

  ltbl <- kable(tbl, digits=2, format="latex", booktabs=TRUE, escape=FALSE,
                linesep="",
                align=(rep("c",ncol(tbl))))  %>%
    column_spec(column=ncol(tbl)+1, border_left=TRUE)
  cat(ltbl, file=paste(rootdir,
                       sprintf("tex/tables_and_figures/dieff-cases-%s.tex",ifelse(use.national, "NI","SI")),
                       sep="/"))

  kable(tbl)
}
```

## Deaths (with deaths as information)

```{r deathtablesd, eval = TRUE, cache=FALSE, warning=FALSE, results='asis'}
df$zero <- 0
infovarsd <- list(c("dlogdd", "logdd"),
                  c("dlogdd", "logdd","dlogdd.national", "logdd.national"))
# infovarsd <- list(c("logdd")
L <- 21
sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
#sdf <- subset(df, df$date>=as.Date("2020-02-29")) #+L)
regs <- mainregressions(sdf,
                        "dlogdd", pols, bvars, infovarsd, "zero", xlist, ilist, iv, L=L)
showhtmltables(regs$pib[[1]], regs$pbiy, regs$piy);
savetextables(regs$pib[[1]], NULL, NULL, prefix="behavior-deathsinfo")
savetextables(c(regs$pib[[1]][1:4], regs$pib[[2]][1:4]), NULL, NULL, prefix="behavior-deathsinfo-NI")
savetextables(c(regs$pib[[1]][1:4]), NULL, NULL, prefix="behavior-deathsinfo-stateinfo")
savetextables(c(regs$pib[[2]][1:4]), NULL, NULL, prefix="behavior-deathsinfo-nationinfo")
savetextables(NULL, regs$pbiy, regs$piy, prefix="deaths-deathsinfo")
savetextables(NULL, regs$pbiy[c(1,3)], regs$piy[c(1,3)], prefix="deaths-deathsinfo-nosplit")
```

```{r dieff-deaths, eval = FALSE, cache=FALSE, results='asis', warning=FALSE}
for (use.national in c(TRUE,FALSE)) {
  sb <- ifelse(use.national, 2, 1)
  sy <- ifelse(use.national, 3, 1)
  pib <- lapply(regs$pib[[sb]][1:4], function(x) x$reg)
  pbiy <- regs$pbiy[[sy]]$reg
  piy <- regs$piy[[sy]]$reg

  S <- 999
  bs <- bootstrap_felm(sdf, c(pib,list(pbiy,piy)), S=S)

  diall <- dieff_table(lapply(regs$pib[[sb]][1:4], function(x) x$reg$coef[,1]),
                       regs$pbiy[[sy]]$reg$coef[,1],
                       regs$piy[[sy]]$reg$coef[,1],
                       policies=c(pols, infovarsd[[sb]]),
                       nsum=length(pols)
                       )

  bdiall <- lapply(1:S, function(i)
    dieff_table(lapply(1:4, function(j) bs[[j]][,i]),
                bs[[5]][,i], bs[[6]][,i],
                policies=c(pols, infovarsd[[sb]]),
                nsum=length(pols)
                ))

  tbl <- addse(diall, bdiall)

  #tbl <- tbl[-c(nrow(tbl), nrow(tbl)-1),]

  ltbl <- kable(tbl, digits=2, format="latex", booktabs=TRUE, escape=FALSE,
                linesep="",
                align=(rep("c",ncol(tbl))))  %>%
    column_spec(column=ncol(tbl)+1, border_left=TRUE)
  cat(ltbl, file=paste(rootdir,
                       sprintf("tex/tables_and_figures/dieff-deaths-deathsinfo-%s.tex",ifelse(use.national, "NI","SI")),
                       sep="/"))
  kable(tbl)
}
```

<!-- ## $\log(\Delta D/\Delta C)$ -->

<!-- ```{r logdddc, cache=FALSE, warning=FALSE, results='asis'} -->
<!-- df$logdddc <- log(df$ddeath/panellag(df$cases, df$state, df$date, 7)) -->
<!-- df$logdddc[df$ddeath<=0] <- NA -->
<!-- df$logdddc[is.infinite(df$logdddc)] <- NA -->
<!-- df$logtestrate <- log(df$testrate) -->
<!-- df$logtestrate[df$testrate<=0] <- NA -->
<!-- regs <- mainregressions(df, "logdddc", pols, bvars, infovars, "lag(logtestrate,7)", xlist[1], ilist[1], list("0"), L=21) -->

<!-- showhtmltables(NULL, regs$pbiy, regs$piy); -->
<!-- ``` -->

## Robustness Check
```{r regsrobust, cache=FALSE, warning=FALSE, results='asis'}
infovars <- list(c("dlogdc", "logdc"),
                 c("dlogdc", "logdc","dlogdc.national", "logdc.national")) #,
#c("dlogdc", "logdc"))
#infovars2 <- list(c( "logdc"))
tvars <- "dlogtests"
interactions <-list("month", statevars)
ilist <- list(interactions, interactions)
iv <- list("0","0") #,"(testratedc ~ lag(testratedc,14))")

# L <- 14
#   xlist <- list("","")
#   plist <- list(pols, pols)
#   ijs <- expand.grid(1:length(bvars), 1:length(plist))
#   pib <- list()
#   for (k in 1:length(xlist)) {
#     pib[[k]] <- apply(ijs, 1, function(ij) {
#       policyreg(df, bvars[ij[1]], NULL,  bvars,
#                 c(plist[[ij[2]]],infovars[[k]],
#                   xlist[[k]]),
#                 ilist[[k]], "0", L=14)
#     })
#   }

# pi-y with log(vote)
xlist <-list(c("logvote"), c("logvote"))
L <- 14
sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
regs <- mainregressions(sdf,
                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
showhtmltables(NULL, NULL, NULL, regs$ip)
savetextables(NULL, NULL, NULL, prefix="policy",ip = regs$ip[1:6])

L <- 14
sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
piy <- list()
yvar <- "dlogdc"
for (k in 1:length(infovars)) {
  # (1) policy index
  pols <- c("pmaskbus","pk12","pindex")
   xlist <-list("","")
  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
  piy[[(1+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
              c(sprintf("lag(%s, %d)", infovars[[k]], L),
                tvars, xlist[[k]]),
              ilist[[k]], 0, L=L)
  # (2) excluding NY
  pols <- c("pmaskbus","pk12","pmovie","pshelter","prestaurant","pnonessential")
   xlist <-list("","")
  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
  sdf2 <- sdf[which(sdf$state!='New York'),]
  piy[[(2+6*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
              c(sprintf("lag(%s, %d)", infovars[[k]], L),
                tvars, xlist[[k]]),
              ilist[[k]], 0, L=L)
  # (3) adding mask wearing rate
  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
  xlist <-list(c("mask_percent"), c("mask_percent"))
  piy[[(3+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
              c(sprintf("lag(%s, %d)", infovars[[k]], L),
                tvars, xlist[[k]]),
              ilist[[k]], 0, L=L)
  # (4) adding log(Trump's vote)
  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
  xlist <-list(c("logvote"), c("logvote"))
  piy[[(4+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
              c(sprintf("lag(%s, %d)", infovars[[k]], L),
                tvars, xlist[[k]]),
              ilist[[k]], 0, L=L)
  # (5) lagged behavior variables
  sdf <- subset(df, df$date>=as.Date("2020-02-22")) #+L)
  xlist <-list("","")
  piy[[(5+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
              c(sprintf("lag(%s, %d)", bvars, L+14),
                sprintf("lag(%s, %d)", infovars[[k]], L),
                tvars, xlist[[k]]),
              ilist[[k]], 0, L=L)
  # (6) (2)-(5)
   sdf <- subset(df, df$date>=as.Date("2020-02-22")) #+L)
   sdf2 <- sdf[which(sdf$state!='New York'),]
  xlist <-list(c("mask_percent","logvote"), c("mask_percent","logvote"))
   piy[[(6+6*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
              c(sprintf("lag(%s, %d)", bvars, L+14),
                sprintf("lag(%s, %d)", infovars[[k]], L),
                tvars, xlist[[k]]),
              ilist[[k]], 0, L=L)
}

colnames(piy[[1]]$reg$response) <- "dlogdc"
showhtmltables(NULL, NULL, piy);
savetextables(NULL, NULL,  piy[1:6], prefix="cases-robust")

L <- 21
sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
piy <- list()
yvar <- "dlogdd"
infovars <- list(c("dlogdd", "logdd"),
                  c("dlogdd", "logdd","dlogdd.national", "logdd.national"))
#xlist <-list(c("z.mask","logvote"), c("z.mask","logvote"))
for (k in 1:length(infovars)) {

  # (1) policy index
  pols <- c("pmaskbus","pk12","pindex")
  xlist <-list("","")
  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
  piy[[(1+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
              c(sprintf("lag(%s, %d)", infovars[[k]], L),
                 xlist[[k]]),
              ilist[[k]], 0, L=L)
  # (2) excluding NY
  pols <- c("pmaskbus","pk12","pmovie","pshelter","prestaurant","pnonessential")
  xlist <-list("","")
  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
  sdf2 <- sdf[which(sdf$state!='New York'),]
  piy[[(2+6*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
              c(sprintf("lag(%s, %d)", infovars[[k]], L),
                 xlist[[k]]),
              ilist[[k]], 0, L=L)

  # (3) adding mask wearing rate
  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
  xlist <-list(c("mask_percent"), c("mask_percent"))
  piy[[(3+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
              c(sprintf("lag(%s, %d)", infovars[[k]], L),
                 xlist[[k]]),
              ilist[[k]], 0, L=L)
  # (4) adding log(Trump's vote)
  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
  xlist <-list(c("logvote"), c("logvote"))
  piy[[(4+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
              c(sprintf("lag(%s, %d)", infovars[[k]], L),
                 xlist[[k]]),
              ilist[[k]], 0, L=L)
  # (5) lagged behavior variables
  sdf <- subset(df, df$date>=as.Date("2020-02-22")) #+L)
  xlist <-list("","")
  piy[[(5+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
              c(sprintf("lag(%s, %d)", bvars, L+14),
                sprintf("lag(%s, %d)", infovars[[k]], L),
                 xlist[[k]]),
              ilist[[k]], 0, L=L)
  # (6) (2)-(5)
   sdf <- subset(df, df$date>=as.Date("2020-02-22")) #+L)
   # sdf2 <- sdf[which(sdf$state!='New York' & sdf$state!='New Jersey' & sdf$state!='Connecticut' & sdf$state!='Massachusetts'),]
   sdf2 <- sdf[which(sdf$state!='New York'),]
  xlist <-list(c("mask_percent","logvote"), c("mask_percent","logvote"))
   piy[[(6+6*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
              c(sprintf("lag(%s, %d)", bvars, L+14),
                sprintf("lag(%s, %d)", infovars[[k]], L),
                xlist[[k]]),
              ilist[[k]], 0, L=L)

}
colnames(piy[[1]]$reg$response) <- "dlogdd"
showhtmltables(NULL, NULL, piy);
savetextables(NULL, NULL,  piy[1:6], prefix="deaths-robust")

```

## Fixed Effects
```{r regsfe, cache=FALSE, warning=FALSE, results='asis'}

source(paste(rootdir,"cases_and_policies/rmd/debiasedfe.R", sep="/"))

df$pindex <- (df$pshelter+df$pmovie+df$prestaurant+df$pnonessential)/4
#df$pindex <- (df$pshelter+df$prestaurant+df$pnonessential)/3

yvar <- "dlogdc"
infovars <- list(c("dlogdc", "logdc"),
                 c("dlogdc", "logdc","dlogdc.national", "logdc.national"))
tvars <- "dlogtests"
L <- 14
sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
piy <- list()

#' Update the coefficients of felm object.  Also updates the t stats
#' and p values, so that summary and stargazer work correctly after.
replacecoef <- function(reg, newcoef) {
  reg$beta[] <- newcoef
  reg$coefficients[] <- newcoef
  reg$tval <- reg$beta/reg$se
  reg$pval <- pt(-abs(reg$tval),df=reg$df)*2
  reg$rtval <- reg$beta/reg$rse
  reg$rpval <- pt(-abs(reg$rtval),df=reg$df)*2
  reg$ctval <- reg$beta/reg$cse
  reg$cpval <- pt(-abs(reg$ctval),df=reg$df)*2
  return(reg)
}

for (k in 1:length(infovars)) {
  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
  pols <- c("pmaskbus","pk12","pmovie","pshelter","prestaurant","pnonessential")
  m <-  reg_fe(sdf, yvar, pols, NULL,
              c(sprintf("lag(%s, %d)", infovars[[k]], L),
                tvars), 0, L, "state + month")
  piy[[(1+4*(k-1))]] <- m
  m$reg <- replacecoef(m$reg, m$bc2)
  piy[[(2+4*(k-1))]] <- m
  pols <- c("pmaskbus","pk12","pindex")
  m <-  reg_fe(sdf, yvar, pols, NULL,
              c(sprintf("lag(%s, %d)", infovars[[k]], L),
                tvars), 0, L, "state + month")
  piy[[(3+4*(k-1))]] <- m
  m$reg <- replacecoef(m$reg, m$bc2)
  piy[[(4+4*(k-1))]] <- m
}

showhtmltables(NULL, NULL,  piy);
prefix="cases-fe"
#savetextables(NULL, NULL, piy[1:4], prefix="cases-fe") copying and

# pasting this is bad style and seems like it might create maintenance
# problems ...
   ylbl <- colnames(piy[[1]]$reg$response)
    if (ylbl=="dlogdd") ylbl <- "Death Growth"
    if (ylbl=="dlogdc") ylbl <- "Case Growth"

    tbl <- capture.output(stargazer(
        lapply(piy[1:4], function(x) x$reg),
        type="latex",
        dep.var.labels.include=FALSE,
        title=ylbl,
        # omit=omit,
        # omit.labels=omit.labels,
        column.labels=c(yvar),
        column.separate=c(length(piy[1:4])),
        # add.lines=list(c("$\\sum_j \\mathrm{Policy}_j$",
        #                  sapply(piy, function(x) printstars(x$peff[1], x$peff[2]))),
        #                c("",
        #                  sprintf("(%.3f)",sapply(piy, function(x) x$peff[2])))
        #                ),
        omit.stat=c("f", "ser"), model.names=FALSE,
        no.space=TRUE,
        column.sep.width="1pt",
        df=FALSE, header=FALSE,
        model.numbers=TRUE))
    tbl <- relabel(tbl)
    tbl <- gsub("Model (\\d)", "\\(\\1\\)", tbl)
    #tbl <- gsub("No","Yes", tbl)
    texfile <- sprintf("%s/tex/tables_and_figures/%s-piy.tex",rootdir,prefix)
    cat(paste(tbl[c(-1,-2,-3,-4, -length(tbl))], collapse="\n"), file=texfile)

yvar <- "dlogdd"
infovars <- list(c("dlogdd", "logdd"),
                  c("dlogdd", "logdd","dlogdd.national", "logdd.national"))
L <- 21
sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
piy <- list()
for (k in 1:length(infovars)) {
  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
  pols <- c("pmaskbus","pk12","pmovie","pshelter","prestaurant","pnonessential")
  m <-  reg_fe(sdf, yvar, pols, NULL,
              sprintf("lag(%s, %d)", infovars[[k]], L), 0, L, "state + month")
  piy[[(1+4*(k-1))]] <- m
  m$reg <- replacecoef(m$reg, m$bc2)
  piy[[(2+4*(k-1))]] <- m
  pols <- c("pmaskbus","pk12","pindex")
  m <-  reg_fe(sdf, yvar, pols, NULL,
              sprintf("lag(%s, %d)", infovars[[k]], L), 0, L, "state + month")
  piy[[(3+4*(k-1))]] <- m
  m$reg <- replacecoef(m$reg, m$bc2)
  piy[[(4+4*(k-1))]] <- m
}

colnames(piy[[1]]$reg$response) <- "dlogdd"
showhtmltables(NULL, NULL,  piy);
colnames(piy[[1]]$reg$response) <- "dlogdd"
showhtmltables(NULL, NULL,  piy);
#savetextables(NULL, NULL, piy[1:4], prefix="deaths-fe")
prefix <- "deaths-fe"
   ylbl <- colnames(piy[[1]]$reg$response)
    if (ylbl=="dlogdd") ylbl <- "Death Growth"
    if (ylbl=="dlogdc") ylbl <- "Case Growth"

    tbl <- capture.output(stargazer(
        lapply(piy[1:4], function(x) x$reg),
        type="latex",
        dep.var.labels.include=FALSE,
        title=ylbl,
        # omit=omit,
        # omit.labels=omit.labels,
        column.labels=c(yvar),
        column.separate=c(length(piy[1:4])),
        # add.lines=list(c("$\\sum_j \\mathrm{Policy}_j$",
        #                  sapply(piy, function(x) printstars(x$peff[1], x$peff[2]))),
        #                c("",
        #                  sprintf("(%.3f)",sapply(piy, function(x) x$peff[2])))
        #                ),
        omit.stat=c("f", "ser"), model.names=FALSE,
        no.space=TRUE,
        column.sep.width="1pt",
        df=FALSE, header=FALSE,
        model.numbers=TRUE))
    tbl <- relabel(tbl)
    tbl <- gsub("Model (\\d)", "\\(\\1\\)", tbl)
    #tbl <- gsub("No","Yes", tbl)
    texfile <- sprintf("%s/tex/tables_and_figures/%s-piy.tex",rootdir,prefix)
    cat(paste(tbl[c(-1,-2,-3,-4, -length(tbl))], collapse="\n"), file=texfile)
```

## DML

```{r regsdml, cache=FALSE, warning=FALSE, results='asis'}
library(glmnet)

df$logpop <- log(df$Population.2018)
df$logsq <- log(df$Square.Miles)
df$repub <- as.factor(df$party)
df$party_dum <- ifelse(df$party == "republican", 0, 1)
df$month3 <- ifelse(df$month==3,0,1)
df$month4 <- ifelse(df$month==4,0,1)
df$month5 <- ifelse(df$month==5,0,1)
df$month6 <- ifelse(df$month==6,0,1)
L <- 14
df$dlogdc.L14 <- panellag(df$dlogdc,df$state,df$date, lag=L)
df$logdc.L14 <- panellag(df$logdc,df$state,df$date, lag=L)
df$dlogdd.L14 <- panellag(df$dlogdd,df$state,df$date, lag=L)
df$logdd.L14 <- panellag(df$logdd,df$state,df$date, lag=L)
df$pmaskbus.L14 <- panellag(df$pmaskbus,df$state,df$date, lag=L)
df$pk12.L14 <- panellag(df$pk12,df$state,df$date, lag=L)
df$pmovie.L14 <- panellag(df$pmovie,df$state,df$date, lag=L)
df$pshelter.L14 <- panellag(df$pshelter,df$state,df$date, lag=L)
df$prestaurant.L14 <- panellag(df$prestaurant,df$state,df$date, lag=L)
df$pnonessential.L14 <- panellag(df$pnonessential,df$state,df$date, lag=L)
df$pindex.L14 <- panellag(df$pindex,df$state,df$date, lag=L)

xvars <- c("logpop",
               "logsq",
               "Percent.Unemployed..2018.",
               "Percent.living.under.the.federal.poverty.line..2018.",
               "Percent.at.risk.for.serious.illness.due.to.COVID","party_dum",
               "mask_percent", "logvote",
               "month3","month4","month5","month6","dlogdc.L14","logdc.L14")

pols.L14 <- c("pmaskbus.L14","pk12.L14","pmovie.L14","pshelter.L14","prestaurant.L14","pnonessential.L14")

sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)

DML2.for.PLM <- function(x, d, y, dreg, yreg, sdf, nfold=5) {
  unit <- as.double(sdf$state)
  j <-  quantile(unit, c(1:(nfold-1))/nfold)
  I <- vector(mode = "list", length = nfold)
  I[[1]] <- unit <= j[1]
  for (b in 2:(nfold-1)) {
    I[[b]] <- (unit <= j[b])&(unit > j[b-1])
  }
  I[[nfold]] <- unit > j[nfold-1]
  nobs <- nrow(x)
  ytil <- dtil <- rep(NA, nobs)
  for(b in 1:nfold){
    I1 <- (I[[b]]==FALSE)
    I2 <- I[[b]]
    dfit <- dreg(x[I1,], d[I1]) # take a fold out
    yfit <- yreg(x[I1,], y[I1]) # take a folot out
    dhat <- predict(dfit, x[I2,], type="response") #predict the fold out
    yhat <- predict(yfit, x[I2,], type="response") #predict the fold out
    dtil[I2] <- (d[I2] - dhat) #record residual
    ytil[I2] <- (y[I2] - yhat) #record residual
    cat(b," ")
  }
  sdf$ytil <- ytil
  sdf$dtil <- dtil
  rfit <- felm(ytil ~ dtil| 0| 0 | state, data=sdf) #estimate the main parameter by regressing one residual on the other
  coef.est <- coef(rfit)[2] #extract coefficient
  se <- sqrt(vcov(rfit)[2,2])
  cat(sprintf("\ncoef (se) = %g (%g)\n", coef.est , se))
  return( list(coef.est = coef.est , se=se) )
}

# dreg <- function(x,d){ rlasso(x, d, c=2, POST=FALSE) } #ML method = rlasso
# yreg <- function(x,y){ rlasso(x, y, c=2, POST=FALSE) } #ML method = rlasso
# DML2.lasso <- DML2.for.PLM(x, d, y, dreg, yreg, sdf)

x = as.matrix(cbind(sdf[,pols.L14[-1]],sdf[,xvars]))
d = sdf$pmaskbus.L14
y = sdf$dlogdc

dreg <- function(x,d){ cv.glmnet(x, d) } #ML method= rlasso
yreg <- function(x,y){ cv.glmnet(x, y) } #ML method = rlasso
set.seed(1)
DML2.lasso <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold = 5)
set.seed(1)
DML2.lasso <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold = 10)

dreg <- function(x,d){ randomForest(x, d) } #ML method=Forest
yreg <- function(x,y){ randomForest(x, y) } #ML method=Forest
set.seed(1)
DML2.RF <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold=5)
set.seed(1)
DML2.RF <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold=10)

x = as.matrix(cbind(sdf[,pols.L14[-4]],sdf[,xvars]))
d = sdf$pshelter.L14
y = sdf$dlogdc

dreg <- function(x,d){ cv.glmnet(x, d) } #ML method= rlasso
yreg <- function(x,y){ cv.glmnet(x, y) } #ML method = rlasso
set.seed(1)
DML2.lasso <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold = 5)
set.seed(1)
DML2.lasso <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold = 10)

dreg <- function(x,d){ randomForest(x, d) } #ML method=Forest
yreg <- function(x,y){ randomForest(x, y) } #ML method=Forest
set.seed(1)
DML2.RF <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold=5)
set.seed(1)
DML2.RF <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold=10)

x = as.matrix(cbind(sdf[,pols.L14[1:2]],sdf[,xvars]))
d = sdf$pindex.L14
y = sdf$dlogdc

dreg <- function(x,d){ cv.glmnet(x, d) } #ML method= rlasso
yreg <- function(x,y){ cv.glmnet(x, y) } #ML method = rlasso
set.seed(1)
DML2.lasso <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold = 5)
set.seed(1)
DML2.lasso <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold = 10)

dreg <- function(x,d){ randomForest(x, d) } #ML method=Forest
yreg <- function(x,y){ randomForest(x, y) } #ML method=Forest
set.seed(1)
DML2.RF <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold=5)
set.seed(1)
DML2.RF <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold=10)

xvars <- c("logpop",
               "logsq",
               "Percent.Unemployed..2018.",
               "Percent.living.under.the.federal.poverty.line..2018.",
               "Percent.at.risk.for.serious.illness.due.to.COVID","party_dum",
               "mask_percent", "logvote",
               "month3","month4","month5","month6","dlogdd.L14","logdd.L14")

y = sdf$dlogdd
dreg <- function(x,d){ cv.glmnet(x, d) } #ML method= rlasso
yreg <- function(x,y){ cv.glmnet(x, y) } #ML method = rlasso
set.seed(1)
DML2.lasso <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold = 5)
set.seed(1)
DML2.lasso <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold = 10)

dreg <- function(x,d){ randomForest(x, d) } #ML method=Forest
yreg <- function(x,y){ randomForest(x, y) } #ML method=Forest
set.seed(1)
DML2.RF <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold=5)
set.seed(1)
DML2.RF <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold=10)

x = as.matrix(cbind(sdf[,pols.L14[-4]],sdf[,xvars]))
d = sdf$pshelter.L14

dreg <- function(x,d){ cv.glmnet(x, d) } #ML method= rlasso
yreg <- function(x,y){ cv.glmnet(x, y) } #ML method = rlasso
set.seed(1)
DML2.lasso <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold = 5)
set.seed(1)
DML2.lasso <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold = 10)

dreg <- function(x,d){ randomForest(x, d) } #ML method=Forest
yreg <- function(x,y){ randomForest(x, y) } #ML method=Forest
set.seed(1)
DML2.RF <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold=5)
set.seed(1)
DML2.RF <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold=10)

x = as.matrix(cbind(sdf[,pols.L14[1:2]],sdf[,xvars]))
d = sdf$pindex.L14
y = sdf$dlogdd

dreg <- function(x,d){ cv.glmnet(x, d) } #ML method= rlasso
yreg <- function(x,y){ cv.glmnet(x, y) } #ML method = rlasso
set.seed(1)
DML2.lasso <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold = 5)
set.seed(1)
DML2.lasso <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold = 10)

dreg <- function(x,d){ randomForest(x, d) } #ML method=Forest
yreg <- function(x,y){ randomForest(x, y) } #ML method=Forest
set.seed(1)
DML2.RF <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold=5)
set.seed(1)
DML2.RF <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold=10)
```

# Figures

## Policy

The following plot illustrates the effect of mandatory masks for
employees.  The lighter line is the average across states that did not
have mask policy 14 days ago of case growth rate. The darker line is
the average across states that did have a mask policy 14 days
ago. (It's actually the fit from regressing dlogdc on date dummies
interacted with the 14 day lagged, 7 day moving average of a mask
policy indicator). Light points are the state observations without a
mask policy. Dark points are states with a mask policy 14 days ago.

```{r fig, cache=FALSE,  fig_width=10, fig_height=6, warning=FALSE}
pfigure <- function(df, pol, stlab=1, outcome="dlogdc", L=14) {
  df$p <- panellag(df[,pol],df$state, df$date, L)
  df$p <- ceiling(df$p)
  df$y <- df[,outcome]
  df$week <- week(df$date)
  m <- lm(y ~ as.factor(date) + as.factor(date):p, data=df)
  adf <- rbind(data.frame(date=m$xlevels$`as.factor(date)`, p=0),
               data.frame(date=m$xlevels$`as.factor(date)`, p=1))
  adf$week <- week(as.Date(adf$date))
  pred <- predict(m, newdata=adf, interval="confidence")
  adf$hat <- pred[,"fit"]
  adf$lo <- pred[,"lwr"]
  adf$hi <- pred[,"upr"]
  adf$date <- as.Date(adf$date)
  if (stlab==1) {
    pdf <- subset(df, df$p<=0.01)
    ldf <- subset(df, df$p>0.01)
  } else if (stlab==0) {
    pdf <- subset(df, df$p>=0.99)
    ldf <- subset(df, df$p<0.99)
  } else {
    pdf <- df
    ldf <- NULL #data.frame(date=NA,y=NA,ST=NA,p=NA)
  }
  clrs <- solarized_pal(2)(2)
  fig <- ggplot(df,aes(color=p)) +
    geom_point(data=pdf,aes(x=date, y=y, color=p),alpha=0.3, size=0.5) +
    xlim(c(as.Date("2020-04-10"), max(df$date))) +
    geom_line(data=adf, aes(x=date, y=hat, group=p), size=2) +
    figtheme + scale_color_gradient(low=clrs[1], high=clrs[2]) + #colors_grad(palette="Green-Gold") +
    theme(legend.position="none") +
    xlab("") +
    ylim(c(-1.2,1.2))
  if (!is.null(ldf))
    fig <- fig +
      geom_text(data=ldf,
                aes(x=date, y=y, label=ST, color=p), alpha=1.0, size=5)
  return(fig)
}
fig <- pfigure(df, "pmaskbus", stlab=-1) + ggtitle(TeX(relabel("dlogdc given pmaskbus"))) + ylab(TeX(relabel("dlogdc")))

p <- "pmaskbus"
figdev(sprintf("%s/tex/tables_and_figures/%s-cases.pdf", rootdir, p), width=4, height=3)
print(fig)
dev.off()
fig
```

We now look at similar figures for other policies.

```{r polplots, cache=FALSE, fig_width=10, fig_height=6}
fig <- pfigure(df, "pk12", stlab=-1) +
  ggtitle(TeX(relabel("dlogdc given pk12"))) +
  xlim(c(as.Date("2020-03-15"), max(df$date))) +
  ylim(c(-1,3)) + ylab(TeX(relabel("dlogdc")))

p <- "pk12"
figdev(sprintf("%s/tex/tables_and_figures/%s-cases.pdf", rootdir, p), width=4, height=3)
print(fig)
dev.off()
fig

for (p in pols[3:length(pols)]) {
  fig <- pfigure(df, p, stlab=-1) +
    ggtitle(TeX(relabel(sprintf("dlogdc given %s",p)))) +
    xlim(c(as.Date("2020-04-01"), max(df$date))) +
    ylim(c(-1,1)) + ylab(TeX(relabel("dlogdc")))
  figdev(sprintf("%s/tex/tables_and_figures/%s-cases.pdf", rootdir, p), width=4, height=3)
  print(fig)
  dev.off()
}
```

# Death Growth Plots

```{r poldeathsplots, cache=FALSE, fig_width=10, fig_height=6, warning=FALSE}
fig <- pfigure(df, "pmaskbus", stlab=-1, outcome="dlogdd", L=21) +
  ggtitle(TeX(relabel("dlogdd given pmaskbus"))) +
  ylab(TeX(relabel("dlogdd")))

p <- "pmaskbus"
figdev(sprintf("%s/tex/tables_and_figures/%s-deaths.pdf", rootdir, p),width=4, height=3)
print(fig)
dev.off()
fig

# fig <- pfigure(df, "pmaskbus", stlab=-1, outcome="dlogdd", L=14) +
#   ggtitle(TeX(relabel("dlogdd given pmaskbus"))) +
#   ylab(TeX(relabel("dlogdd")))
# p <- "pmaskbus"
# figdev(sprintf("%s/tex/tables_and_figures/%s-deaths.pdf", rootdir, p),width=4, height=3)
# print(fig)
# dev.off()
# fig

fig <- pfigure(df, "pk12", stlab=-1, outcome="dlogdd", L=21) +
  ggtitle(TeX(relabel("dlogdd given pk12"))) +
  xlim(c(as.Date("2020-03-15"), max(df$date))) +
  ylim(c(-1,3)) + ylab(TeX(relabel("dlogdd")))

p <- "pk12"
figdev(sprintf("%s/tex/tables_and_figures/%s-deaths.pdf", rootdir, p),width=4, height=3)
print(fig)
dev.off()

for (p in pols[3:length(pols)]) {
  fig <- pfigure(df, p, stlab=-1, outcome="dlogdd", L=21) +
    ggtitle(TeX(relabel(sprintf("dlogdd given %s",p)))) +
    xlim(c(as.Date("2020-04-01"), max(df$date))) +
    ylim(c(-1,1)) + ylab(TeX(relabel("dlogdd")))
  figdev(sprintf("%s/tex/tables_and_figures/%s-deaths.pdf", rootdir, p), width=4, height=3)
  print(fig)
  dev.off()
}
```

```{r, results='asis', echo=FALSE}
if (!knitr::is_latex_output()) {
  cat("
[![](https://i.creativecommons.org/l/by-sa/4.0/88x31.png)](http://creativecommons.org/licenses/by-sa/4.0/)

This work is licensed under a [Creative Commons Attribution-ShareAlike
4.0 International License](http://creativecommons.org/licenses/by-sa/4.0/)
") }
```