### 2020-09-01 cases_and_policies/rmd/cf-pindex.Rmd
Fixed titles of mask figures.

---
title       : "Causal impact of masks, policies, behavior on early COVID-19 pandemic in the U.S."
subtitle    : "Counterfactuals"
author      :  Victor Chernozhukov,  Hiro Kasahara, Paul Schrimpf
job         :
date        : "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: "covid.bib"
link-citations: true
always_allow_html: true
output      :
    html_document :
        toc : true
        toc_depth : 3
        toc_float : true
        number_sections : true
        #theme : journal
        #css : 628notes.css
        code_folding: hide
        #lib_dir : deps
        self_contained : true
        fig_width: 8
        fig_height: 6
---

 <!-- To create html files from this, in R enter  -->
 <!-- library(rmarkdown) -->
 <!-- rmarkdown::render("regressionsWithPlots.Rmd") -->
 <!-- To create tex files from this, in R enter ` -->
 <!-- library(rmarkdown) -->
 <!-- rmarkdown::render("regressionsWithPlots.Rmd", output_format=latex_fragment()) -->

 <!-- to create the gh-pages hosted website in R enter -->
 <!-- rmarkdown::render_site() -->

 <!-- Then in a shell  -->
 <!-- $ ghp-import -p -n -m "$(date)" _site -->

```{r, include=FALSE}
options(knitr.table.format = function() {
  if (knitr::is_latex_output()) 'latex' else 'pandoc'
})
if (knitr::is_latex_output()) {
  knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo=FALSE)
}
```

# Data Preparation
```{r setup, cache=FALSE, warning=FALSE, include=FALSE}

library(lfe)
library(stargazer)
library(knitr)
library(plm)
library(latex2exp)
library(ggplot2)
library(ggthemes)
library(estimatr)
library(gridExtra)
library(grid)
library(mvtnorm)
library(kableExtra)
library(data.table)
library(reshape2)

sgtype <- opts_knit$get("rmarkdown.pandoc.to")
sgstyle <- 'default'
colors <-  scale_color_solarized
colors_fill <- scale_fill_solarized
rootdir <- system("git rev-parse --show-toplevel", intern=TRUE)[1]
source(paste(rootdir,"cases_and_policies/R/dataprep.R",sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/varlabels.R", sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/regprep.R", sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/generatetables.R", sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/bootstrap_felm.R", sep="/"))

datelims <- c(as.Date("2020-03-07"), max(df$date))
```

```{r cffuncs, cache=FALSE}
source(paste(rootdir,"cases_and_policies/rmd/counterfactuals.R", sep="/"))
pols <- c("pmaskbus","pk12","pshelter","pindex")
```

```{r testsmooth, cache=FALSE}
source("smoothtests.R")
```

```{r regressions, cache=FALSE}
infovars <- list(c("dlogdc", "logdc")) #,
#c("dlogdc", "logdc"))
#infovars2 <- list(c( "logdc"))
tvars <- "dlogtests"
xlist <-list("")
interactions <-list("month", statevars)
ilist <- list(interactions) #, interactions)
iv <- list("0") #,"(testratedc ~ lag(testratedc,14))")

L <- 14
sdf <- subset(df, df$date>=datelims[1]) #+L)
regs <- mainregressions2(sdf,
                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
m <- regs$piy[[1]]$reg
cnames <- c(pols, infovars[[1]])

di <- dieff_table2(lapply(regs$pib[[1]][1:4], function(x) x$reg$coef[,1]),
                  regs$pbiy[[1]]$reg$coef[,1],
                  regs$piy[[1]]$reg$coef[,1],
                  policies=cnames)

pib <- lapply(regs$pib[[1]][1:4], function(x) x$reg)
pbiy <- regs$pbiy[[1]]$reg
piy <- regs$piy[[1]]$reg

S <- 999
bs <- bootstrap_felm2(sdf, c(pib,list(pbiy,piy)), S=S)
bdi <- lapply(1:S, function(i)
  dieff_table2(lapply(1:4, function(j) bs[[j]][,i]),
               bs[[5]][,i], bs[[6]][,i],
               policies=cnames)
  )

subcoefs <- function(di, coef) {
  for (r in 1:(nrow(di)-1)) {
    i <- grep(rownames(di)[r] , rownames(coef))
    coef[i,1] <- di[r,"Average"]
  }
  return(coef)
}

bcoefs <- t(do.call(cbind, lapply(1:S, function(i) {
  coef <- matrix(bs[[6]][,i],ncol=1)
  rownames(coef) <- rownames(bs[[6]])
  return(subcoefs(bdi[[i]],coef ))
})))

V <- cov(bcoefs)
m$coefficients <- subcoefs(di, m$coefficients)
m$vcv <- V
m$clustervcv <- V
m$robustvcv <- V
m$beta <- m$coefficients
m$cse <- sqrt(diag(V))

rerunsim <- FALSE
```

# Method

We compute counterfactuals from the "total effect" version of the model, with behavior concentrated out.

$$
Y_{i,t+\ell}
= \mathsf{a}'
P_{it} + \mathsf{b}' I_{it} +  {\tilde {\delta}}' W_{it}   + {\bar \varepsilon}_{it}
$$

We consider a counterfactual change of $P_{it}$ to $P_{it}^*$, while
$W_{it}$ and $\bar \varepsilon_{it}$ are held constant. In response to
the policy change, $Y_{i,t+\ell}$ and the part of $I_{it}$ that
contains $Y_{it}$, change to $Y_{i,t+\ell}^*$ and $I_{it}^*$. To be
specific, let $C_{it}$ denote cumulative cases in state $i$ on day
$t$. Our outcome is:

$$
Y_{i,t+\ell} \equiv \log
\left(C_{i, t+\ell} - C_{i, t + \ell - 7} \right) - \log
\left(C_{i, t+\ell-7} - C_{i, t + \ell - 14} \right) \equiv \Delta
\log \Delta C_{i,t+\ell}
$$

where $\Delta$ is a 7 day difference operator. Writing the model in
terms of $\Delta \log \Delta C$ we have:

$$
\Delta \log \Delta C_{i,t+\ell} = \mathsf{a}' P_{it} + \mathsf{b}_D
\Delta \log \Delta C_{i,t} + \mathsf{b}_L \log \Delta C_{i,t} +
{\tilde {\delta}}' W_{it}   + {\bar \varepsilon}_{it}
$$

To simplify computation, we rewrite this model in terms of $\log
\Delta C$:

$$
\log \Delta C_{i,t+\ell} = \mathsf{a}' P_{it} + \log \Delta
C_{i,t+\ell-7} + (\mathsf{b}_D + \mathsf{b}_L) \log \Delta C_{i,t} -
\mathsf{b}_D \log \Delta C_{i,t-7} + {\tilde {\delta}}' W_{it}   + {\bar \varepsilon}_{it}
$$

This equation is used to iteratively compute
$\log \Delta C_{i,t+\ell}^*$
conditional on initial $\{\log \Delta C_{i,s}\}_{s=-\ell-7}^{0}$,
and the entire sequence of $\{W_{i,t}, P_{i,t}^*, \bar \varepsilon_{it} \}_{t=0}^T$.

Weekly cases instead of log weekly cases are given simply by

$$ \Delta C_{i,t+\ell}^* = \exp(\log \Delta C_{i,t+\ell}^*)$$

Cumulative cases can be recursively computed as:

$$
  C_{i,t} = C_{i,t-7} + \Delta C_{i,t} = C_{i,t-7} + \exp(\log \Delta C_{i,t})
$$

given initial $\{C_{i,s}\}_{s=-7}^0$.

Note that $\log \Delta C_{i,t+\ell}^*$ depends linearly on
$\bar \varepsilon$, so the residuals (and our decision to condition on
versus intergrate out) do not matter when considering linear
constrasts of $\log \Delta C_{it}$ or $\Delta \log \Delta C_{it}$, or
when considering relative constrasts of $\Delta C_{it}$.

On the other hand, cumulative cases are a nonlinear function of
$\bar\varepsilon$, so the conditioning on $\bar\varepsilon$ might
matter (not that there's anything wrong with that).

## Inference

Let $S(\theta, \mathbf{\varepsilon})$ denote some counterfactual
quantity or contrast of interest, where $\theta = (\mathsf{a},
\mathsf{b}, \tilde{\delta})$ are the parameters, and
$\mathbf{\varepsilon}$ is the vector of residuals. Examples of $S$
that we compute include:

- Contrasts of growth rates: $S(\theta, \mathbf{\varepsilon}) =
  \Delta \log \Delta C_{it}^* - \Delta \log \Delta C_{it}$

- Relative contrasts of weekly cases: $S(\theta, \mathbf{\varepsilon}) =
  \Delta C_{it}^*/\Delta C_{it}$

- Constrasts of cumulative cases: $S(\theta, \mathbf{\varepsilon}) =
  C_{it}^*/C_{it}$

The first two examples do not actually depend on
$\mathbf{\varepsilon}$, but the third one does. Inference is by
simulation. Let $\hat{\theta}$ denote our point estimates and
$\hat{\mathbf{\varepsilon}}$ the associated residuals. We draw
$\tilde{\theta}_j$ from the asymptotic distribution of
$\hat{\theta}$. Currently the code then computes

$$ \tilde{s}_j = S(\tilde{\theta}_j, \hat{\mathbf{\varepsilon}}) $$

for $j=1,..., 200$ and plots the mean across $j$ as a point estimate
and quantiles across $j$ for confidence intervals. This is fine when
$\varepsilon$ does not affect the quantity of interest, but I think
the correct thing to when $\varepsilon$ matters is to compute

$$ \tilde{s}_j = S(\tilde{\theta}_j, \tilde{\mathbf{\varepsilon}}_j) $$

where $\tilde{\mathbf{\varepsilon}}_j$ are the residuals from using
$\tilde{\theta}_j$ as coefficients.

# Cases

## Masks

We now look at what happens when all states impose a mask mandate on
March 14th.

```{r mask_sim, cache=FALSE}
sdf <- subset(df, df$date>=datelims[1]) #+L)
simmodel <- createforwardsim(m, sdf)
abbrv <- unique(df[,c("state","ST")])

cf <- df
cf[,"Mandate.face.mask.use.by.employees.in.public.facing.businesses"]  <- as.Date("2020-03-14")
cf[,"pmaskbus"] <-
  smoothedpolicy(cf,"Mandate.face.mask.use.by.employees.in.public.facing.businesses",type="ma",bw=6)
cf <- subset(cf, cf$date>=datelims[1]) #+L)
maskmodel <- createforwardsim(m, cf)

if (!file.exists("cfmask_idx.Rda") || rerunsim) {
  cfmasklist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=maskmodel, data=sdf))
  maskall <- rbindlist(lapply(cfmasklist, getmeans))
  save(cfmasklist, maskall, file="cfmask_idx.Rda")
}
load("cfmask_idx.Rda")
cfmask <- do.call(rbind, lapply(cfmasklist, function(x) x$edf))
```

```{r, cache=FALSE}
cfmask <- merge(cfmask, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfmask, data=df)
  print(f$figp +  ggtitle(sprintf("Effect of mandating masks on March 14th in %s",st)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dgrowth_idx.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dcases_idx.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dcumu_idx.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdcum)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-rcumu_idx.pdf",rootdir,st), width=5,height=3.75)
  print(f$figrcum)
  dev.off()

  if (st=="Washington") for (fig in f) print(fig)
}

figs <- nationalplots(maskall, cfmask, "mandating masks on March 14th\n")

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dgrowth_idx.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dcases_idx.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-rel_idx.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-rcumu_idx.pdf",sep="/"), width=5,height=3.75)
figs$cr
dev.off()

for (f in figs) print(f)
```

## Stay at home

We now look at what happens if stay at home orders were never issued.

```{r home, cache=FALSE}
cf <- df
cf[,"pshelter"] <- 0
cf <- subset(cf, cf$date>=datelims[1]) #+L)
noshelter <- createforwardsim(m, cf)

if (!file.exists("cfshelter_idx.Rda") || rerunsim) {
  cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf))
  shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
  save(cfshelterlist, shelterall, file="cfshelter_idx.Rda")
}
load("cfshelter_idx.Rda")
cfshelter <- do.call(rbind, lapply(cfshelterlist, function(x) x$edf))
```

```{r, cache=TRUE}
cfshelter <- merge(cfshelter, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfshelter, data=df)
  print(f$figp +  ggtitle(sprintf("Effect of no stay at home order in %s",st)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dgrowth_idx.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dcases_idx.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dcumu_idx.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdcum)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-rcumu_idx.pdf",rootdir,st), width=5,height=3.75)
  print(f$figrcum)
  dev.off()

}

figs <- nationalplots(shelterall, cfshelter, "no stay at home orders")

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dgrowth_idx.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dcases_idx.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-rel_idx.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-rcumu_idx.pdf",sep="/"), width=5,height=3.75)
figs$cr
dev.off()

for (f in figs) print(f)
```

## Averaged policies

We now look at what happens if none of the three averaged policies
(close movie theaters, close restaurants, close
nonessential businesses) were implemented.

```{r index, cache=FALSE}
cf <- df
cf[,"pindex"] <- 0
cf <- subset(cf, cf$date>=datelims[1]) #+L)
noindex <- createforwardsim(m, cf)

if (!file.exists("cfindex_idx.Rda") || rerunsim) {
  cfindexlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noindex, data=sdf))
  indexall <- rbindlist(lapply(cfindexlist, getmeans))
  save(cfindexlist, indexall, file="cfindex_idx.Rda")
}
load("cfindex_idx.Rda")
cfindex <- do.call(rbind, lapply(cfindexlist, function(x) x$edf))
```

```{r, cache=TRUE}
cfindex <- merge(cfindex, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfindex, data=df)
  print(f$figp +  ggtitle(sprintf("Effect of no shutdown policies in %s",st)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-index-dgrowth_idx.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-index-dcases_idx.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-index-dcumu_idx.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdcum)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-index-rcumu_idx.pdf",rootdir,st), width=5,height=3.75)
  print(f$figrcum)
  dev.off()

}

figs <- nationalplots(indexall, cfindex, "no shutdown policies")

figdev(paste(rootdir,"tex/tables_and_figures/us-index-dgrowth_idx.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-index-dcases_idx.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-index-rel_idx.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-index-rcumu_idx.pdf",sep="/"), width=5,height=3.75)
figs$cr
dev.off()

for (f in figs) print(f)
```

# Deaths

```{r statesim-deaths, cache=FALSE}
infovars <- list(c("dlogdd", "logdd")) #,
#c("dlogdc", "logdc"))
#infovars2 <- list(c( "logdc"))
tvars <- NULL
xlist <-list("")
interactions <-list("month", statevars)
ilist <- list(interactions) #, interactions)
iv <- list("0") #,"(testratedc ~ lag(testratedc,14))")

L <- 21
sdf <- subset(df, df$date>=datelims[1]) #+L)
regs <- mainregressions2(sdf,
                        "dlogdd", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)

m <- regs$piy[[1]]$reg

cnames <- c(pols, infovars[[1]])
di <- dieff_table2(lapply(regs$pib[[1]][1:4], function(x) x$reg$coef[,1]),
                  regs$pbiy[[1]]$reg$coef[,1],
                  regs$piy[[1]]$reg$coef[,1],
                  policies=cnames)

pib <- lapply(regs$pib[[1]][1:4], function(x) x$reg)
pbiy <- regs$pbiy[[1]]$reg
piy <- regs$piy[[1]]$reg

S <- 999
bs <- bootstrap_felm2(sdf, c(pib,list(pbiy,piy)), S=S)
bdi <- lapply(1:S, function(i)
  dieff_table2(lapply(1:4, function(j) bs[[j]][,i]),
              bs[[5]][,i], bs[[6]][,i],
              policies=cnames)
  )

bcoefs <- t(do.call(cbind, lapply(1:S, function(i) {
  coef <- matrix(bs[[6]][,i],ncol=1)
  rownames(coef) <- rownames(bs[[6]])
  return(subcoefs(bdi[[i]],coef ))
})))
V <- cov(bcoefs)
m$coefficients <- subcoefs(di, m$coefficients)
m$beta <- m$coefficients
m$vcv <- V
m$clustervcv <- V
m$robustvcv <- V
m$cse <- sqrt(diag(V))

simmodel <- createforwardsim(m, sdf, dvar="dlogdd", var="logdd")

# remove policies
cf <- sdf
cf[,pols] <- 0
nopmodel <- createforwardsim(m, cf, dvar="dlogdd", var="logdd")

rerunsim=FALSE
```

## Masks

We now look at what happens when all states impose a mask mandate on
March 14th.

```{r mask_sim-deaths, cache=FALSE}
cf <- df
cf[,"Mandate.face.mask.use.by.employees.in.public.facing.businesses"]  <- as.Date("2020-03-14")
cf[,"pmaskbus"] <-
  smoothedpolicy(cf,"Mandate.face.mask.use.by.employees.in.public.facing.businesses",type="ma",bw=6)
cf <- subset(cf, cf$date>=datelims[1]) #+L)
maskmodel <- createforwardsim(m, cf, dvar="dlogdd", var="logdd")

if (!file.exists("cfmask_deaths_idx.Rda") || rerunsim) {
  cfmasklist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=maskmodel, data=sdf, var="logdd", dvar="dlogdd"))
  maskall <- rbindlist(lapply(cfmasklist, getmeans))
  save(cfmasklist, maskall, file="cfmask_deaths_idx.Rda")
}
load("cfmask_deaths_idx.Rda")
cfmask <- do.call(rbind, lapply(cfmasklist, function(x) x$edf))
```

```{r, cache=TRUE}
cfmask <- merge(cfmask, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfmask, yvar="deaths", data=df)
  print(f$figp +  ggtitle(sprintf("Effect of mandating masks on March 14th in %s",st)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dgrowth_deaths_idx.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dcases_deaths_idx.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dcumu_deaths_idx.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdcum)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-rcumu_deaths_idx.pdf",rootdir,st), width=5,height=3.75)
  print(f$figrcum)
  dev.off()

}

figs <- nationalplots(maskall, cfmask, "mandating masks on March 14th\n", yvar="deaths")

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dgrowth_deaths_idx.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dcases_deaths_idx.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-rel_deaths_idx.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-rcumu_deaths_idx.pdf",sep="/"), width=5,height=3.75)
figs$cr
dev.off()

for (f in figs) print(f)
```

## Stay at home

We now look at what happens if stay at home orders were never issued.

```{r home-deaths, cache=FALSE}
cf <- df
cf[,"pshelter"] <- 0
cf <- subset(cf, cf$date>=datelims[1]) #+L)
noshelter <-  createforwardsim(m, cf, dvar="dlogdd", var="logdd")

if (!file.exists("cfshelter_deaths_idx.Rda") || rerunsim) {
  cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf))
  shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
  save(cfshelterlist, shelterall, file="cfshelter_deaths_idx.Rda")
}
load("cfshelter_deaths_idx.Rda")
cfshelter <- do.call(rbind, lapply(cfshelterlist, function(x) x$edf))
```

```{r, cache=TRUE}
cfshelter <- merge(cfshelter, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfshelter, data=df, yvar="deaths")
  print(f$figp +  ggtitle(sprintf("Effect of no stay at home order in %s",st)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dgrowth_deaths_idx.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dcases_deaths_idx.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dcumu_deaths_idx.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdcum)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-rcumu_deaths_idx.pdf",rootdir,st), width=5,height=3.75)
  print(f$figrcum)
  dev.off()

}

figs <- nationalplots(shelterall, cfshelter, "no stay at home orders",yvar="deaths")

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dgrowth_deaths_idx.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dcases_deaths_idx.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-rel_deaths_idx.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-rcumu_deaths_idx.pdf",sep="/"), width=5,height=3.75)
figs$cr
dev.off()

for (f in figs) print(f)
```

## Shutdown policies

We now look at what happens if the three business related shutdown policies in pindex were never implemented.

```{r index-deaths, cache=FALSE}
cf <- df
cf[,"pindex"] <- 0
cf <- subset(cf, cf$date>=datelims[1]) #+L)
noindex <- createforwardsim(m, cf, dvar="dlogdd", var="logdd")

if (!file.exists("cfindex_deaths_idx.Rda") || rerunsim) {
  cfindexlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noindex, data=sdf))
  indexall <- rbindlist(lapply(cfindexlist, getmeans))
  save(cfindexlist, indexall, file="cfindex_deaths_idx.Rda")
}
load("cfindex_deaths_idx.Rda")
cfindex <- do.call(rbind, lapply(cfindexlist, function(x) x$edf))
```

```{r, cache=TRUE}
cfindex <- merge(cfindex, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfindex, data=df, yvar="deaths")
  print(f$figp +  ggtitle(sprintf("Effect of no shutdown policies in %s",st)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-index-dgrowth_deaths_idx.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-index-dcases_deaths_idx.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-index-dcumu_deaths_idx.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdcum)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-index-rcumu_deaths_idx.pdf",rootdir,st), width=5,height=3.75)
  print(f$figrcum)
  dev.off()

}

figs <- nationalplots(indexall, cfindex, "no shutdown policies", yvar="deaths")

figdev(paste(rootdir,"tex/tables_and_figures/us-index-dgrowth_deaths_idx.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-index-dcases_deaths_idx.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-index-rel_deaths_idx.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-index-rcumu_deaths_idx.pdf",sep="/"), width=5,height=3.75)
figs$cr
dev.off()

for (f in figs) print(f)
```